\documentclass{article}


\usepackage[margin=1.2in]{geometry}
\usepackage{parskip}

% \input{pdg-preamble-v1}
\input{confidence-preamble}

\addbibresource{conf.bib}

% \newcommand{\ext}[1]{\mathcal M#1} %measures over phi
% \newcommand{\ext}[1]{\mathbb R^{#1}_{\mathrm{fin}}} % finite measures over phi
\newcommand{\ext}[1]{\overline #1} %  measures over phi
\newcommand{\Unif}{\mathrm{Unif}}

\def\cofunc{commitment function}
\def\confdom{\mathdcal C}

\DeclareMathOperator{\supp}{\mathrm{Supp}}
\DeclareMathOperator{\Tr}{Tr}


\begin{document}

\begin{center}
    \Large\scshape
    % Confidence: How to think of partial updates
    % Confidence, the
    Updating with Variable Confidence
    % Partial Updates
    % (Belief Opacity?)
\end{center}
\bigskip

\tableofcontents

% You have confidence in a piece of information.
% The word ``confidence'' is sometimes used to stand in for a piece of information.
% While high confidence in an event is much like high probability in the event, low confidence in the event and low probability are entriely different.


\section{Introduction}
\def\stmt{$A$}
% \def\stmt{$\phi$}

% The ability express information with varying confidence is an important aspect of knowledge representation.
% Being able to articulate things with variable confide

The ability articulate a ``degree of confidence'' is an important aspect of knowledge representation.
% We will present the concept 
% We give a formal account of a new concept
We are about to propose a concept . 
% Subpoint: it helps avoid a brittleness of always beliving things
% Subpoint: Protects against overconfidence.
Now, there are many well-established ways of representing and measuring uncertainty,
probability chief among them. 
Indeed, an informal poll of our colleagues suggests that computer scientists view ``confidence'' as a synonym for probability. 
The enormous success of probability seems to have shadowed the concept we have in mind, which is closely related, but fundementally different. 

% , such as belief functions, and most importantly, probabilities.
% In each case, one adopts a more complex belief state, 
% The enormous success of probability in particular has had an enormous impact on the way computer scientists talk about 
% The success of probability has shadowed other concepts that might well also be called measures of confidence.
% Indeed, it 

% There are actually two related, but slightly different notions here. 
% % The first is a measurement of
% There is a second setting in which one might interpret the word confidence, as well, in an updating context: it's how seriously you take an input 
% Note the difference: 





Confidence is also a scale between two extremes. It is a measure of trust on a scale of completley untrustworthy $(\bot)$ to fully trusted $\top$. 
High confidence is quite like high probability: if we really trust a statement \stmt, we should fully incorporate it into our beliefs, and thereby come to believe it with high probability. 
Similarly, it only makes sense to be extremely confident in a \stmt\ if you believe that \stmt\ is extremely likely to be true. 
Low confidence, on the other hand, is quite different from low probability. 
If we have little trust in \stmt, we should \emph{ignore} \stmt, rather than coming to believe that \stmt\ is unlikely.
% To say \stmt\ has low probabilty is  high confidence in the $\lnot$\stmt. 
For example, if an adversary tells you something that you happen to already believe, you may have low confidence in their statement, but nevertheless ascribe it high probability. 
Importantly, zero confidence represents a truly neutral stance; a statement with zero confidence has no effect.  

We will analyze this notion of confidence in the context of updating.
In this setting, one has some belief state, and recieves inputs, which one might have some degree of confidence in, which is used to modify one's belief state. 
Confidence measures how seriously to take an input in updating beliefs. 
% The approach we take in this paper is very general, 
% Our approach is quite general, and applies any time a belief state is modified in resonse to an input. 
% one wants to measure (gradual) incorporation of new information into one's beliefs, and may be thought of as a measure of how much of the new information makes it into one's beliefs. 


\subsection{Relations to Other Concepts}

\textbf{Probability.}
Probability is a numerical scale that ranges from untenable (0) to undeniable (1). 
No number on this scale is truly neutral.
%  probability of $\frac12$ .
$\frac12$ may split the difference between the extremes, but is by no means always a neutral assessment: learning that your favored candidate is likely to win with probabitliy $\frac12$ is a big deal, if a win was previously thought to be inevitable. 
This shortcoming has perhaps been the primary selling point of many alternatives to probabiltiy, such as Dempster-Shafer Belief functions. 


\textbf{Shafer's Weight of Evidence.}
Shafer's ``weight of evidence'' is precisely the same concept we have in mind. 
Our analysis precsely reduces to his, in the setting where belief states are Belief functions over a set $W$, and observations are subsets of $W$. 
This paper generalizes these concepts to a wider variety of settings, so that one might use one's own 


\textbf{Confidence Intervals and Variance.}
Variance is also commonly used to measure confidence. 
If a sensor is unreliable and can give a range of answers, the variance of the estimate is a very common way of quantifying this reliablility.
If measurements have zero variance, in some sense one has absolute confidence ($\top$) in the sensor. If measurements have infinite variance, then in some sense one has no confidence in the sensor, since individual samples convey no information about the true value of the quantity measured.
In \cref{sec:variance}, we will see how, in some settings, confidence and inverse variance coincide.

% For example, suppose we recieve an input \stmt.  If we completely trust \stmt, we should integrate it fullly into our beliefs.  Thus, the effect of having high confidence in \stmt\ is that we come to believe \stmt\ with high probability.
% If we know that the input comes from an untrustworthy source, for instance, the best thing to do is to ignore it entirely, even if 

% because we want to be able to deal with information that we aren't certain about.

%
% It is extremely useful to be able to express variable confidence in information
%%
%%
%%
% Among computer scientists, the notions of confidence and probability are often identified.
% Among computer scientists, ``confidence'' is often a synonym for probability.
% % ---%
% % Although the two notions are closely related, it is important to distinguish between confidence and likelihood.
% For instance, it is common to say that one has ``high confidence in a statement \stmt'' to indicate a belief that
% % (the content of) \stmt is likely to be true.
% \stmt\ is likely to be true.
% But
% % However,
% having high confidence in \stmt\ can also mean something subtly different: that we should \emph{trust} \stmt, in the sense of taking it seriously in updating our beliefs.
% % The two notions might seem indistinghishable at first,
% % This distinction may not at first seem meaningful, because if we take \stmt\ seriously in updating our beliefs, we will ultimately believe \stmt\ is likely.
% If we take \stmt\ seriously in updating our beliefs, we will ultimately believe \stmt\ is likely, and so the distinction might not seem meaningful at first.
% % However, the two are quite different at the other end of the spectrum. If you believe that \stmt\ is
% % However, thinking that one should \emph{not} trust \stmt\ is quite different from thinking that \stmt\ is unlikely.
% % But \emph{low} probability is quite different from low probability.
% % At the other end of the scale, however, the two notions differ a great deal.
% But at the other end of the scale, the two notions are quite different.
% If we have little trust in \stmt, we should \emph{ignore} \stmt, rather than coming to believe that \stmt\ is unlikely.
% So, if an untrusted party tells you something that you happen to already believe, you might reasonably ascribe it high probability but low confidence.

% We argue that confidence is a useful concept, distinct from probability.
% We will show how confidences
% In this paper, we give a formal framework for talking about confidence in the context of updating,
% which we then use to interperet quantities across a wide variety of settings as confidences.
% % In each case, we will see the following features:
% Here are some recurring themes:

\subsection{Recurring Themes}

\begin{enumerate}[nosep]
    \item Updating with zero confidence leaves the belief state unchanged.
    \item Updating with absolute certainty (the limit as confidence becomes large) causes you to (permanently) gain a certainty, and acts like a projection.
    % \item A zero-confidence update does nothing, while
    % \item Zero confidence is neutral.
    % updating with certainty (the limit of large confidence) is a projection.
    % \item Updating with certainty (the limit of large confidence) is a projection.
    \item Confidence measures involve an arbitrary choice of scale, and so only ratios of confidences are meaningful.
    % \item $\exp(- \beta E)$
    % \item Confidence tends to end up as the scaling term for an exponential functions
    % \item Update rules are exponential in confidence.
\end{enumerate}


% Theorem. Certainties can never be gained or lost with any finite sequence of differentiable updates.
% Theorem. Certainty is always infinitely far away, as measured by the Riemannian metric.

\section{The Opacity Analogy}

% \begin{minipage}{0.48\linewidth}

    So far as a computer display is concerned, what ultimately matters are pixel colors.
    And yet, in the standard representation of a color color, only three of the four bytes actually encodes color information; the last one describes opacity---in some sense, a \emph{confidence} in the color.
    % Far from an oversight, this enables the possibility of suggest a color with less than complete confidence is incredibly useful.
    % The result is an ability to suggest a color with less than complete confidence.
    This enables a great deal in graphics: intricate digital art, via semantically meaningful layers,
        visual effects such as fire and fog that do not need to be modeled jointly with their surroundings.
    % This is not an oversight; working with translucency can be incredibly valuable.
     % porgrams enables .%
        % \footnote{just as higher-level programming languages enable programmers to build larger projects by giving larger-scale structure}



% \end{minipage}
% \hfill
% \begin{minipage}{0.48\linewidth}
    So far as statistical models are concerned, what ultimately matters are probabilities of events.
    But that doesn't appear to


\section{A Formalism for Updating}
\def\X{\mathcal X}
\def\W{\mathcal W}

The setup is as follows.
You have some belief state $\theta \in \Theta$. 
% $\Theta$ might be \textellipsis
% \begin{itemize}[nosep]
%     \item $\Delta W$, the set of probability distributions over measurable space $W$ of outcomes;
%     \item the set of parameters to some parametric family of distributions;
%     \item the set of belief functions over a common space of outcomes;
%     \item the set of all possible PDGs;
%     \item a set of worlds an agent considers possible;
%     \item 
% \end{itemize}
It might be 
a probability distribution, a parameter setting to some parameteric family of probabilities, a set of probabilities, a Dempster-Shafer belief function, or even just a setting of weights for a neural network. 
You recieve some input information $\phi \in \Phi$, such as an event or a sample from a dataset, which you may use to update your belief state. 

Before we add confidence into the mix, let's first see how this setup works without it.
The primary assumption that we have to make is:
\begin{CFaxioms}
    \item[\textbf{F.}] 
        The updating process can be described as a function $F : \Theta \times \Phi \to \Theta$, which produces a new belief state, as a function of the old belief state and the new information.
\end{CFaxioms}
In curried form, $F : \Phi \to (\Theta \to \Theta)$. 

% We now proceed with the formal details.
\textbf{Update Rules.}
Consider a space $\Theta$
of possible belief states,
and a set $\Phi$ of statements.
% and a set $\Phi$ of ``statements'', i.e., the things one can have confidence in.
% An \emph{update rule} (or more precisely, a \emph{$\Theta$-updating rule on $\Phi$})
An \emph{update rule}, or more precisely, a \emph{$\Theta$-update rule on $\Phi$},
is a function of the form
\[
    % F :  (\mathbb R \times \Phi) \to \Big( \Theta \to \Theta \Big)
    F :  \Phi \to \Big( \Theta \to \Theta \Big)
\]
% which describes how to update beliefs about $X$, with the new information, at a certain level of trust.
which describes how to (fully) update beliefs $\Theta$ with new information $\Phi$.
Given $F$ and a statement $\phi$, we call $F_\phi : \Theta \to \Theta$ an \emph{update function},
and for $F$ to be an update rule, we require that $F_\phi$ is
% require that it be idempotent.
idempotent ($F_\phi \circ F_\phi = F_\phi$), meaning that updating any belief with $\phi$ twice in a row is equivalent to single update.
Intuitively, this is because if we have just ({fully}) updated our beliefs to be consistent with the information $\phi$, then a second observation of $\phi$ will require no further alterations of our belief state.
% Having said that, one reading of this paper is a relaxation of this requirement.
Here are some examples.
\begin{enumerate}
    \item %\textbf{Conditioning.}
    \textbf{Conditioning.}
    Let $W = \{w_1, \ldots, w_n\}$ be a finite set of ``possible worlds'',
    $\Theta := \Delta W$ be the set of probability distributions over $W$,
    and $\Phi := 2^W$ be set of all subsets of $W$.
    One particularly natural $\Delta W$-update rule on $2^W$ is given by conditioning,
    \[
    \begin{aligned}
        (-) \smash{\,\Big|\,} (\;\cdot\;) : \qquad 2^W &\to (\Delta W \to \Delta W) \\
        % (-) \,\Big|\, (\;\cdot\;) : \Delta X \times 2^X \to \Delta X.
        % (\mu \in \Delta X, A \subset X) &\mapsto \mu | A .
        (A \subset W) &\mapsto (  ~~\mu~~ \mapsto \mu \mid A ),
    \end{aligned}
    % \qquad
    % \qquad
    % \begin{tikzpicture}[center base]
    %     \node[dpad0] (W) {$W$};
    %     \node[dpad0, right=0.5 of W] (Aq) {$A?$};
    %     \draw[arr2, <-] (W) to node[above]{$\mu$} ++(-1, 0);
    %     \draw[arr2, ->>] (W) to node[above,pos=0.3]{} (Aq);
    %     \draw[arr2, <<-] (Aq) to node[above,pos=0.7]{$A!$} ++(1.2,0);
    % \end{tikzpicture}
    \]
    % where $(\mu \mid A)(x) = \frac{\mu(\{x\})}{\mu(A)}$
    % in which learning $A$ maps
    % where the action of the conditional measure $\mu\mid A$ is given by $(\mu \mid A) \{w\} = \ifrac{\mu\{w\}}{\mu(A)}$.
    where the action of the conditional measure $\mu\mid A$ is given by $(\mu \mid A)(B) = \ifrac{\mu(B \cap A)}{\mu(A)}$, provided $\mu(A) > 0$,
    and can be defined arbitrarily otherwise.
    Of course, it is not the only $\Delta W$-updating rule for $2^W$.
     % $\Phi$.

    \item
    \textbf{Imaging.}
    A second example of an update rule is the ``imaging'' approach of David Lewis
    \parencite{lewis1976probabilities}.
    % , albeit in very different notation.
    Once again, consider a finite set $W$, and belief states $\Theta := \Delta W$.
    For any set $\Phi$, suppose that we already have an $W$-update rule
    $f : \Phi \to (W \to W)$, which we interpret as assigning, to each statement $\phi \in \Phi$ and $w \in W$, a unique world $f_\phi w \in W$ which is the world ``most similar to $w$, in which $\phi$ is true''.
    In this case, idempotence of $f_\phi$ amounts to the very reasonable requirement that the world most similar to $f_\phi w$ in which $\phi$ is true, is $f_\phi w$ itself.
    From $f$, we can construct a $\Delta W$-update rule by
    \[
        F_\phi(\mu) :=
            f^{\sharp}(\mu)
            =
            A \mapsto \mu( f^{-1}_\phi( A ))
        \qquad
        \qquad
        \begin{tikzpicture}[center base]
            \node[dpad0] (W) {$W$};
            \node[dpad0, right=1 of W] (W') {$W$};
            \node[dpad0, below right=0.2 and 0.2 of W] (Phi) {$\Phi$};
            \mergearr{W}{Phi}{W'}
            \node[above=1pt of center-WPhiW']{$f$};
            \draw[arr2, <-] (W) to node[above]{$\mu$} ++(-1, 0);
            \draw[arr2, <<-] (Phi) to node[below]{$\phi$} ++(-1.3, 0);
            \draw[arr2, <-, dashed, gray] (W') to node[above]{$F_\phi(\mu)$} ++(2, 0);
        \end{tikzpicture}
    \]
    which is the pushforward measure of $\mu$ through $f_\phi$, which Lewis calls the ``image of $\mu$ on $\phi$'', and intuitively acts by moving all of the probability mass on the world $w$ to the ``closest world'' $f_\phi w$.
    Since $f$ is idempotent, $F$ will be as well.


    \commentout{
    \item More generally, consider a measurable space $\W = (W, \mathcal A)$, where $W$ is a set and $\mathcal A$ is a $\sigma$-algebra over $W$, and let $\mathcal F \subset \mathcal A$ be closed under supersets in $\mathcal A$.
    % Now, let $\Theta$ be the set of conditional probabili$

    \TODO[Properly Use Conditional Probability Measure, to define on all events]

    Conditioning a probability distribution $\mu \in \Delta\X$ on an event $A \in \mathcal A$ also makes sense in this more general measure-theoretic setting, at least so long as $\mu(A) > 0$, and is given by
    % the Lebesgue integral
    % \[
    $$
        % (\mu \mid A) (B) = \frac{1}{\mu(A)} \int \mathbf 1_{B}(x)  \mathrm d\mu(x)
        (\mu \mid A) (B) = \frac{\mu(B \cap A)}{\mu(A)}
    $$
    }


    \item \textbf{Jeffrey's Rule.}
    Once more, suppose that $W$ is a finite set and $\Theta := \Delta W$. But now our observations will not be measurable events (subsets of $W$), but instead pairs $(X,\pi)$,
    % written $\pi(X)$,
    where $X : W \to S$ is a random variable (i.e., some function of $W$), and $\pi \in \Delta S$ is a distribution over the possible values that $X$ can take.
    Jeffrey's rule, given by
    \begin{align*}
        \mathrm{Jeffrey}_{(X,\pi)} (\mu) &:= \sum_{x \in S} \pi(X\!=x) \;  \mu \,\Big|\, \{ w : X(w) = x \}
            \\
            &= A \mapsto \sum_{x \in S} \pi(X\!=x)\, \mu( A \mid X \!= x)
    \end{align*}

    When $\pi(X) = \delta_x$ is a point mass $X=x$, then Jeffrey's Rule simply conditions on the event $X = x$.
     % but for other choices of $\pi(X)$,
    For this reason, Jeffrey's Rule is sometimes seen as a way of making an update with uncertainty (i.e., a low-confidence update), but as we will see, it instead is rather a high-confidence update on a more expresive class of observations.

    \item \textbf{Conjunction Learning}.
    Suppose $X_1, \ldots, X_n$ are binary variables, and we are trying to learn a conjunction
    $h(X_1, \ldots X_n) = \bigwedge_{j \in J} X_j \bigwedge_{j \in \bar J} \bar X_j$,
    where some of the variables might be negated.
    % and let $T := \{ x_1, \lnot x_1, \ldots, x_n, \lnot x_n\}$ be all possible settings of one of the variables.
    Let $\Theta$ be the set of all such conjunctions, which can be generated by pairs of of subsets of variables $J,\bar J \subset \{1, \ldots, n\}$.
    Fix some ``true'' (but unknown) conjunction $f \in \Theta$ which generates our observations, and let $\Phi$ be the set of possible pairs $(\mat x, b)$, where
    $\mat x = (X_1\!=\!x_1, \ldots, X_n \!=\!x_n)$ is a joint setting of all variables, and $b = f(\mat x)$.

    One reasonable update rule proceeds as follows. Given a hypothesis $h \in \Theta$ represented by $(J, \bar J)$, and an observation $(\mat x \mapsto b)$:
    \[
        F_{(\mat x, b)}(h) :=
        \begin{cases}
            [J, \bar J] & \text{ if } h(\mat x) = b \\
            [J \setminus \{i : x_i = 0\}, \bar J] &\text{ if } h(\mat x) = 0 \text { but } b = 1\\
            [J, \bar J \setminus \{i : x_i = 1\}] &\text{ if } h(\mat x) = 1 \text { but } b = 0\\
    \end{cases}
    \]
    So, if our current hypothesis $h$ correctly predicts the sample, we do nothing. If it predicts that $h(\mat x)$ is false but $f(\mat x)$ is true, then we restrict
    So, if $h(\mat x) = 1$ but $b = 0$, we

    \TODO


    % \item Many other examples of update rules for standard notions of uncertainty can be found in \parencite[Ch. 3]{halpern2017reasoning}.
\end{enumerate}
%

% , at a certain level of trust.
% We are interested in . A \emph{}
% Each of these update rules takes its input
What about less extreme belief alterations, in which we only gradually incorporate information?
Humans certainly don't update our beliefs with update functions.
The prevalance of flash cards as a learning tool, which are meant to be cycled through more than once, witensses this clearly.
Similarly, artificial neural networks are trained with many incremental updates, and cycle through training data more than once.
% Indeed, essentailly all modern machine learning algorithms work by updating parameters little-by-little, rather than fully incorporating input information.
Broadly speaking, there seems to be value in learning the same thing more than once---or, put another way---in updating with varaible confidence.

To that end, we now consider a domain of variable confidence. Formally, we represent the possible degrees of confidence by a group $(\confdom, +, 0)$ with an absorbing element $\top \in \confdom$, such that $\top + c = \top$ for all $c \in \confdom$.
In terms of confidence, we interpret the components of the group as follows.
\begin{itemize}%[]
    \item
    We view the elements of $\confdom$ as possible degrees of confidence.
    \item The group operation $+ : \confdom \times \confdom \to \confdom$ describes how to combine two confidences in the same assertion, to obtain a new confidence that represents the combined confidence in the assertion.
    \item The neutral element $0 \in \confdom$ indicates ``no confidence'' in an observation.
        %
        % The fact that we want to ignore information we have no confidence in
        % gives corresponds to the group identity laws: that
        The group identity laws, which assert that
        $0 + c = c = c + 0$ for all $c \in \confdom$,
        concord with the intuition that an observation that one has no confidence in, has no effect.
        % in which case we should ignore the information at hand,
    \item The absorbing element $\top$, which may be read as ``full confidence'' indicates that we should fully incorporate the information into our belief state.
    The absorbtion property corresponds to the intuition that, definitive information that $\phi$ is true, when combined with other (perhaps less reliable) information that $\phi$ is true, is still definitive.
\end{itemize}

We now define a \emph{\cofunc} to be a function
\[
    F: \confdom \to ( \Phi \to (\Theta \to \Theta))
\]
which, given an incremental confidence $c \in \mathbb \confdom$, returns an ``incremental'' update rule, i.e., a function with the same type as an update rule, except possibly non-idempotent.
Given $c \in \confdom$ and a statement $\phi \in \Phi$, we write
 % Given a piece of information $\phi \in \Phi$, , we write
% $F^\beta_\phi : \Delta\X \to \Delta X$
$F^c_\phi : \Theta \to \Theta$
for the update function prescribed by $F$.
Furthermore, we will insist that update rules obey certain properties.
% , especially the following two.
\begin{CFaxioms}
    % \item \textbf{(zero)} $F^{0}_A(\Pr) = (\Pr)$
    % \item  $F^{0}_A  = {1}_{\Delta\X}$. (That is, $F^{0}_A(\Pr) = \Pr$ for all $\Pr \in \Delta\X$.)
    % \item  $F^{0}_\phi  = {1}_{\Delta\X}$. (That is, $F^{0}_\phi(\Pr) = \Pr$ for all $\Pr \in \Delta\X$.)
    %     \hfill \textbf{(zero)} \label{ax:zero}
    \item  $F^{0}_\phi  = {\mathrm{Id}}_{\Theta}$.
        (That is, $F^{0}_\phi(\theta) = \theta$ for all $\theta \in \Theta$.)
        \hfill \textbf{(zero)} \label{ax:zero}
    % \item $F^{\beta_1}_A \circ F^{\beta_2}_A = F^{\beta_1 + \beta_2}_A$
    % \item $F^{\top} : \Phi \to (\Theta \to \Theta)$ is an update rule, i.e.,
    %     the funciton $F^{\top}_\phi : \Theta \to \Theta$ is an idempotent.
    \item
        % For all $\phi$,
        $F^\top_\phi$ is an idempotent update function.\\
        Equivalently, $F^\top: \Phi \to (\Theta \to \Theta)$ is an update rule.
        % the funciton $F^{\top}_\phi : \Theta \to \Theta$ is an idempotent.
        \hfill \textbf{(idempotence)} \label{ax:idemp}
% \end{CFaxioms}
% The next axiom,
% \begin{CFaxioms}
    % \item For all $\beta_1, \beta_2 \in \mathbb R_{\ge 0}$,~
    \item For all $c_1, c_2 \in \confdom$,~
        $F^{c_1}_\phi \circ F^{c_2}_\phi = F^{c_1 + c_2}_\phi$
        \hfill \textbf{(additivity)} \label{ax:additivity}
\end{CFaxioms}
\Cref{ax:zero} captures the intuition that we should ignore information in which we have no confidence, while \cref{ax:idemp} formalizes the intuition that a full-confidence update should behave as above.
% \Cref{ax:additivity} states that, for all $\theta$, the function $c \mapsto F^c_\phi\theta : \confdom \to \Theta$ is a group homomorphism.
 % a confidence of $\top$ indicates that we are fully incorporating information into our beliefs.
\Cref{ax:additivity} simply states that \cofunc s respect the combination operation: if we fix an assertion $\phi$, then an update with confidence $c_1$ followed by an update with confidence $c_2$ is equivalent to an update with confidence $c_1 + c_2$, which is, by definition, the result of comining confidences $c_1$ and $c_2$.
On its own, so long as we have the freedom to choose $\confdom$, \Cref{ax:additivity} has no teeth.

\begin{prop}
    If $F: \confdom \to (\Phi \to (\Theta \to \Theta))$ satisfies \cref{ax:zero,ax:idemp}, then we can construct a new update
    function for $\Theta$ on $\Phi$, that \emph{is} additive, but with the altered confidence domain
    $$
        \confdom' := \{ \text{ finite lists } [c_1, \ldots c_n]
            \text{ such that } c_i \in \confdom \text{ for all } i = 1, \ldots, n \},
    $$
    whose group operation is list concatenation.
    This update rule is given as follows:
    \[
        AF^{[c_1, \ldots, c_n]}_\phi (\theta)  :=
                (F^{c_n}_\phi \circ \cdots \circ F^{c_1}_\phi) (\theta).
    \]
\end{prop}
\begin{proof}
By construction,
\begin{align*}
AF^{[c_1, \ldots, c_n]}_\phi ( AF^{[d_1, \ldots, d_m]}_\phi (\theta) )
        &:=
            F^{d_m}_\phi \circ \cdots \circ F^{d_1}_\phi (
            F^{c_n}_\phi \circ \cdots \circ F^{c_1}_\phi (\theta))\\
        &= (F^{d_m}_\phi \circ \cdots \circ F^{d_1}_\phi \circ
        F^{c_n}_\phi \circ \cdots \circ F^{c_1}_\phi) (\theta) \\
        &= AF^{[c_1, \ldots, c_n, d_1, \ldots, d_m]}_\phi (\theta) \\
        &= AF^{[c_1, \ldots, c_n] + [d_1, \ldots, d_m]}_\phi (\theta).
\end{align*}
\end{proof}

If $F$ is a \cofunc, then by \cref{ax:idemp}, $F^{\top}$ is an update rule, and we call $F$ a ``refinement'' of the update rule $F^{top}$.

We are primarily interested in the case where confidence can be measured as a real number, and so for most of this paper, we take $\confdom := \mathbb R_{+} = \big([0, \infty) \cup \{\infty\}, +, 0, \infty\big)$ to be the group of extended nonnegative real numbers under addition.
With this choice of confidence domain, \cref{ax:additivity} begins to have more bite, although, as we will see, the effect is more to pin down a coherent system of measurement, and does not appear to restrict modeling expressivness.
%
%
Below is a concrete representative example of a \cofunc\ with our standard confidence domain $\mathbb R_+$.
\begin{examplex}{Balls in Urns}{colorballs}
        %\label{ex:colorballs}
        \def\red{{\mathsf{red}}}
        \def\green{{\mathsf{green}}}
        \def\blue{{\mathsf{blue}}}
        % Perhaps we're uncertain about our shoe size.
    % Imagine a ball has been drawn from an urn,
    % and we are uncertain about its color, which we know to be either red, green, or blue.
    Suppose there is a game show in which, at some point, a ball of unknown color
    (red, green, or blue) is drawn from an urn.
    Let  $X := \{ \red,\green,\blue \}$ be the three-element set corresponding to possible colors,
     and $\mathcal A = 2^X$ be all possible subsets of it, so that a robability $\mu \in \Delta \X$
    can be specified as a
    % three-dimensional vector $\mat p \in \Theta
    point in the simplex
    $$\Theta
    := \Delta \{ \red,\green,\blue \}
    = \{\mat p = (p_1, p_2, p_3)  \in \mathbb R^3_{\ge 0} : p_1 + p_2 + p_3 = 1\},$$
    interpreting
    the coordinates as the respective probabilities of drawing $\red$, $\green$, and $\blue$.
    % the a point
    % $\langle p_0, p_1, p_2 \rangle$
    % in the standard basis as
    % $(p_\red, p_\green, p_\blue \rangle$, where $p_x$ is the probability of drawing color $x$.
    % an interpretation of $\langle p_0, p_1, p_2 \langle$ as the respective probabilities
    %
         % is just a 3-dimensional vector $\mat p$.
    % For the assertion language, consider the set
    Now consider the set
     $\Phi := \{$``the ball is $\mathsf c$''$ : \mathsf c \in X\} \cong X$ of possible assertions of the color of the ball.
    % , so that you can make assertions such as ``the ball is red''.

    % Now, consider the update rule $F$ defined by
    One possible \cofunc\ for $\Phi$ and $\Theta$ can be defined by increasing the probability of the given statement, by a factor exponential in the confidence, and then renormalizing.
    For instance,
    \[
        % F^{\beta}_h( \mat p ) =
        % \frac{1}{\mat p \cdot \mat e_h} \mat p \odot \exp(\beta )
        F^{\beta}_{\text{``the ball is $\red$''}} \left(
            \begin{bmatrix} p_\red \\ p_\green \\ p_\blue \\ \end{bmatrix}
        \right) :=
            \underbrace{\vphantom{\Bigg|}
                \frac{1}{1 + p_\red (e^{\beta} - 1) }
            }_{\text{normalizing constant}}
            % \begin{bmatrix}
            %     \vdots \\
            %     \hphantom{e^\beta} p_{h-1} \\
            %     e^\beta \cdot p_h  \\
            %     \hphantom{e^\beta} p_{h+1}\\
            %     \vdots
            % \end{bmatrix}
            \begin{bmatrix}
                e^\beta  p_\red \\
                p_\green  \\
                p_\blue
            \end{bmatrix},
    \]
    and analogously for green and blue.
    Note that an exponential of some sort is necessary to get \cref{ax:additivity}, for update rules of this kind; we cannot simply replace
    % it with another increasing function of $\beta$.
    $e^{\beta}$ with $\beta^2$, for instance.

    As mentioned above, you can compose udpates. For instance, suppose we had a prior $\mat p_0$ over colors, and then three people come up to us and make assertions about the ball color---%
    % first an aquaintance, then our friend, and then our tailor.
    first someone we don't recognize, who claims the ball is red, then a hobyist with a statistical model who says it's green, and finally someone who says they witnessed the draw, who says blue.
    We trust them to different degrees, and so in response  chain of updates as follows:
    \[
        \mat p_0
            \xmapsto{\displaystyle F^{\,0.3}_{\text{``ball is red''}}}
        \mat p_1
            \xmapsto{\displaystyle F^{\,0.6}_{\text{``ball is green''}}}
        \mat p_2
            \xmapsto{\displaystyle F^{\,7.1}_{\text{``ball is blue''}}}
        \mat p_3.
    \]

    % How should one decide the confidences $\beta$?
    % That depeends on the update rule.

    % The choices of confidence $\beta=0.3, 0.6, 7.1$ seem meaningless, but one aim of this paper is to argue that they are meaningful relative to an update rule.

    % We make the following observations.
    We now make some observations.
    \begin{enumerate}
        % \item The order might matter.
        \item For this particular update rule, the order of the updates does not matter.
        \item If we happen to already know the color of the ball (e.g., if $\mat p_0$ is a point mass on $\red$), then none of these updates will have an effect; this will still be the case after the updates.
        % \item betas are meaningless if we mix and match update rules
        \item Starting from any strictly positive distribution, one can get to any distribution at all.
        Concretely, for every $\mat p$ and $\mat q > 0$, there is a vector
        % $\boldsymbol\beta^{\mat q \to \mat p}
        % = (\beta_{\red}^{\mat q \to \mat p}, \beta_{\green}^{\mat q \to \mat p}, \beta_{\blue}^{\mat q \to \mat p})$
        $\boldsymbol\beta
        = (\beta_{\red}, \beta_{\green}, \beta_{\blue})$
        such that applying the updates
        $F^{\beta_{\red}
            % ^{\mat q \to \mat p}
            }_{\red}$,
        $F^{\beta_{\green}
            % ^{\mat q \to \mat p}
            }_{\green}$, and
        $F^{\beta_{\blue}
            % ^{\mat q \to \mat p}
            }_{\blue}$,
        to $\mat q$, in any order, yields $\mat p$. Moreover, this vector is unique.

        % In general, it will turn out that .

        % By the Radon-Nikodim Theorem, there if $\mu \ll \nu$, there is a measure $\frac{\mathrm d \mu}{\mathrm d\nu}$ such that $\int_{A} \frac{\mathrm d \mu}{\mathrm d\nu} \mathrm d \nu = \mu(A)$
    \end{enumerate}
    \end{examplex}



Here are some more abstract examples of \cofunc s, with confidence domain
$\confdom := \mathbb R_+$.
\begin{enumerate}
\item
Once again, suppose $W$ is a finite set,
$\Theta := \Delta W$, and $\Phi := 2^W$.
Here are two natural \cofunc s for this scenario, both of which are refinements of conditioning.
\begin{itemize}
    \item
    $\displaystyle
        (F1^c_A \mu)(B) = (1-e^{-c}) \mu(B|A) +  e^{-c} \mu(B)
    $
    \item
    $\displaystyle
        (F2^c_A \mu)(B) \propto \mu(B|A)^{(1-e^{-c})} \mu(B)^{e^{-c}}
    $
\end{itemize}
The first \cofunc, $F1$, linearly interpolates between the result of ignoring the information contained in the event $A$ (i.e., leaving the belief state $\mu$ unchanged) and conditioning on $A$.
By contrast, $F2$ does a similar interpolation, but in log space.

\item
    % \textbf{Neural Networks Updates.}
    % \textbf{Machine Learning.}
Suppose that $\Theta$ is the set of possible parameter settings for a neural network, which aims to predict an element of $Y \subset \mathbb R^{m}$ given an in put from $X \in \mathbb R^{n}$.
So, for each $\theta \in \Theta$, we have a function $f_\theta : X \to Y$, and for fixed $x \in X$, the function $\theta \mapsto f_\theta(x) : \Theta$ is differentiable.

% $\{ f_\theta : X \to Y \}_{\theta \in \Theta}$;
% perhaps it is the set of possible weights of a neural network.


 % suppose $\Theta$ is a set of parameter settings ,

% WE DO NOT WANT TO TALK ABOUT DS FUNCTIONS HERE BECAUSE THEY ARE VARIABLE CONFIDENCE.
\item
Again consider a finite set $W$ and suppose $\Theta$ consists of all Dempster-Shafer belief functions
\end{enumerate}


Another confidence domain we could consider is the group
\[
    \mathrm{ZO} := \Big(~ [0,1],
        \quad a \star b := a b,
        \quad 1,
        \quad 0 ~\Big)
\]
Clearly, 1 is neutral and 0 is absorbing element.
\begin{prop}
    $\mathrm{ZO}$ is isomorphic to $\mathbb R_+$, but therere is no canonical choice of isomorphism.
\end{prop}
\begin{proof}
    For every $k > 0$ can construct an isomorphism $\varphi_k: \mathrm{ZO} \to \mathbb R_+$ explicitly by $\varphi(a) := - k \log a$.
    It is a homomorphism, since
    \[
        \varphi(a \star b) = - k \log (a b) = - k \log a - k \log b =
            \varphi(a) + \varphi(b),
    \]
    while $\varphi(1) = 0$ (so it preserves the identity) and $\varphi(0) = \infty$ (so it preserves the absorbing element).
    The inverse mapping can also be explicitly by $\varphi^{-1}(r) := \exp( - r / k)$, which is also a homorphism for the same reasons as above.
\end{proof}

% By currying we
% \begin{prop}
%
% \end{prop}



We are primarily interested in the setting where $\Theta$ parameterizes a family of probaility distributions.
To that end, suppose that $\X = (X, \mathcal A)$ be a measurable space, so that $X$ is a set and $\mathcal A$ be a $\sigma$-algebra over it, let $\Delta \X$ denote the set of probability measures over $\X$,
and keep in the back of our heads an indexed family
% $\{ p_\theta ( X_\theta ) : \theta \in \Theta\}$.
$
    \mathcal P =
    \{ p_\theta \in\Delta\X \mid \theta \in \Theta \}
$ of probability distributions.
% If we take

% For instance, starting with a distribution $\mu_0 \in \Delta \X$, we can \emph{compose} updates.
%
% % \[
% %     \mu_0
% %         \xmapsto{\displaystyle F^{.3}_{\mathit{Height}=5'11''}} \mu_1
% %         \xmapsto{\displaystyle F^{.6}_{\Pr(Y=1|X=3)=0.4}} \mu_2
% %         \xmapsto{\displaystyle F^{2.1}_{K_i(\varphi)}} \mu_3.
% % \]



\subsection{Basic Axioms for Update Rules}

We now detail some other properties we might want an update rule to have.

\textbf{Differentiability}.
% A primary goal of this paper is to study how updates are made in low-confidence settings.
Insofar as confidence is meant to smoothly interpolating between fully incorporating information and ignoring it, we would like the interpolation to be differentiable.

Here are two ways of capturing this, depending on the structure one has in hand.


\begin{CFaxioms}
    \item \label{ax:diffble}
    \begin{enumerate}
    \item If $\Theta$ itself has differentiable structure, we require that
        for fixed $\theta$ and $\phi$, the function $\beta \mapsto F^{\beta}_\phi(\theta)$
        is continuously differentiable. %\label{ax:diffble}
    \item If $\Theta$ parameterizes a family of probabilities over $(\X, \mathcal A)$,
        via $\{ \Pr_\theta \}_{\theta \in \Theta}$, we can avoid talking abot a differentiable structure on $\Theta$ by simply requiring that the update rule be differentiable from the perspective of every event $A \in \mathcal A$.
        That is,
        %  is a family of probability distributions.
        for all $\theta \in \Theta$, $\phi \in \Phi$, and  $A \in \mathcal A$,
        the function $\beta \mapsto \Pr_{F^{\beta}_\phi(\theta)}(A)
        : \mathbb R_{\ge 0} \to \mathbb [0,1]$ is
        continuously differentiable.
        %(in pairs $(\beta,\Pr)$).
            % \hfill \textbf{(differentiability)}
            \label{ax:diffble2}
    \end{enumerate}
    \hfill \textbf{(differentiability)}
\end{CFaxioms}

% Suppose that the space $\Theta$ is actually a differentiable manifold.
% In this case, we might want want $F$ to be compatible with the differentiable structure.
% \begin{CFaxioms}
%     \item $\Theta$ is a differentiable manifold.
%         For fixed $\theta$ and $\phi$, the function $\beta \mapsto F^{\beta}_\phi(\theta)$
%         is continuously differentiable.
%             \hfill \textbf{(differentiability)} \label{ax:diffble2}
% \end{CFaxioms}

\textbf{Certainty Axioms.}
% The simplest thing to do
At the other extreme, we want to describe learning a certainty as the limit of updating with high confidence, so we require that
\begin{CFaxioms}
    % \item[Cert1]
    \item
        % $\lim_{c \to \infty} F^{c}_\phi \in \bar\Theta$ exists, where $\bar\Theta$ is the closure of $\Theta$
        $\displaystyle\lim_{c \to \infty} F^{c}_\phi : \Theta \to \bar\Theta$
        exists, so we may abbreviate it as $F^\infty_\phi$
        \label{ax:cert-exists}
\end{CFaxioms}

It follows from \Cref{ax:additivity,ax:cert-exists} that
% learning a certainty
$F^\infty$
is a ``projection'', in that is an idempotent operation, since,
for all $\theta \in \Theta$,
\[
    F^\infty_\phi (\theta)
        = \lim_{c \to \infty} F^{c}_\phi (\theta)
        % = \lim_{c_1 \to \infty} \lim_{c_2 \to \infty} F^{c_1 + c_2}_\phi (\theta)
        = \lim_{\substack{c_1 \to \infty \\ c_2 \to \infty}}
            F^{c_1 + c_2}_\phi (\theta)
        % = \lim_{c_1 \to \infty} \lim_{c_2 \to \infty}
        = \lim_{\substack{c_1 \to \infty \\ c_2 \to \infty}}
            F^{c_1}_\phi ( F^{c_2}_\phi (\theta))
        % = \lim_{c_2 \to \infty} F^{\infty}_\phi (F^{c_2}_\phi(\theta))
        = \lim_{c_1 \to \infty} F^{c_1}_\phi (F^{\infty}_\phi(\theta))
        % = \lim_{c \to \infty} F^{c}_\phi \,\circ\, F^{c}_\phi %(\theta)
        % F^\infty_\phi
        % = F^{\infty}_\phi (F^{\infty}_\phi(\theta))
        = F^\infty_\phi \,\circ\, F^\infty_\phi  (\theta)
        .
\]

Let us return to the setting where $\Theta$ parameterizes a family of probability distributions over $\X = (X, \mathcal A)$, via a function $\Pr : \Theta \to \Delta\X$,
and suppose that for some assertions $\phi \in \Phi$ represent events $A \in \mathcal A$.
The result of learning such an assertion with certainty should correspond to conditioning.

\begin{CFaxioms}
    \item If $\phi$ represents $A \in \mathcal A$, and $\Pr_{\theta}(A) > 0$, then
        $\displaystyle \lim_{\beta\to\infty} \Pr_{ F^\beta_\phi\theta } = \Pr_\theta \mid A$.
        % \hfill \textbf{(absolute certainty)} \label{ax:certainty}
        \hfill \textbf{(conditioning)} \label{ax:conditioning}

    % \item
    % % $F^\beta_A$ and $F_{\bar A}^\beta$ are inverses.
    % $F^\beta_A \circ F_{\bar A}^\beta = 1_{\Delta\X}$.
    %     \hfill \textbf{(complementation)} \label{ax:comp}
\end{CFaxioms}


We can also consider the weaker variant of \Cref{ax:conditioning}:
\begin{CFaxioms}
    % \item[U3$'$.]  \textbf{(effectiveness)~} $\supp F^\infty_A (\Pr) \subset A $
    % customlabel
    \item[\Cref*{ax:conditioning}$'$]
    % \item[\customlabel{ax:effectiveness}{\textbf{\Cref*{ax:conditioning}$'$.}}]
    If $\phi \in \Phi$ represents $A \in \mathcal A$ and $\Pr(A) > 0$, then
        % $F^\infty_A (\Pr)(A) = 1$
        % $\displaystyle \lim_{\beta\to\infty} F^\beta(\Pr)(A) = 1$
        $\displaystyle \lim_{\beta\to\infty} \Pr_{F^\beta_\phi\theta}(A) = 1$.
        \hfill \textbf{(effectiveness)} \label[CFaxiomsi]{ax:effectiveness}
\end{CFaxioms}
in which the limit of infinite confidence in $A$ is not required be the same as conditioning the starting point on $A$, but still must have the same effect of ensuring that $A$ occurs with probability 1.


\textbf{Symmetry and Commutativity}
Some update rules, such as the one in \cref{ex:colorballs} have a particularly conveneint property: the result of applying many updates does not depend on the order in which the information arives.

\begin{CFaxioms}
    \item For all $\phi_1, \phi_2 \in \Phi$ and $\beta_1, \beta_2 \in \mathbb R$, and
    % $\mu \in \Delta\X$,
    $\theta \in \Theta$,
    we have that
    $
        F^{\beta_1}_{\phi_1} ( F^{\beta_2}_{\phi_2}(\theta)) =
            F^{\beta_2}_{\phi_2} ( F^{\beta_1}_{\phi_1}(\theta)).
    $
    \hfill\textbf{(Commutativity)} \label{ax:commute}
\end{CFaxioms}

We will not always want to insist on commutativity. Human belief updates, for example, are notoriously non-commutative, in part due to confirmation bias:
if you are already fairly certain that $P$ is false, you are likely to disregard
information that $P$ is true. Thus earlier updates tend to be more impactful.%
\footnote{
    It is also possible to model this effect with a commutative update rule,
    by expressing reduced confidence in later inputs.
}


We would also like update rules to preserve any symmetries shared by the state space $\X$ and the assertion language $\Phi$, so that updates are not sensitive to irrelevant labelings of points.
% Concretely, let $\mathrm{Aut}(X, \Phi)$ be the set of automorphisms $\sigma : X \to X$, together with an action on assertions, so that $\sigma\phi \in \Phi$ is the appropriately relabeled assertion equivalent to $\phi$ after the relabeling.
Concretely, let $\mathrm{Aut}(\Theta, \Phi)$ be the set of automorphisms $\sigma : \Theta \to \Theta$ (say, rotations of the simplex of distributions), that also have an associated action on assertions, so that $\sigma\phi \in \Phi$ is the corresponding relabeling of $\phi$ under $\sigma$.  The symmetry condition can now be captured by:

\begin{CFaxioms}
    \item
        % All symmetries of $\X$ are reflected in the update rule.
        % Concretely, for all  $\sigma
        For all $\sigma
            % : X \to X
            % \in \mathrm{Aut}(X, \Phi)$, we have
            \in \mathrm{Aut}(\Theta, \Phi)$, we have
        % \\ \indent\hspace{2em}
        % $F^\beta_A(\sigma_\#(\Pr)) = \sigma_\#\Big(F^\beta_{\{\sigma(a) : a \in A \}}(\Pr)\Big)$,
        % $F^\beta_\phi (\sigma_\#(\Pr)) = \sigma_\#\Big(F^\beta_{\sigma\phi}(\Pr)\Big)$,
        $F^\beta_{\sigma\phi} (\sigma(\theta)) = \sigma \Big( F^\beta_{\phi}(\Pr)\Big)$.
            \hfill \textbf{(symmetry)} \label{ax:symmetry}
        \\
        % where $\sigma_\#(\Pr)$ is the pushforward of $\Pr$ under $\sigma$.
\end{CFaxioms}

Intuitively, this axiom states that doing an update is equivalent to changing to an equivalent representaton, doing the appropriately transformed update, and then transforming back.

% One possible choice of $\Phi$ statements with the $\sigma$-algebra $\mathcal A$ of measurable sets, which are closed under union and complementation, and on which probabilistic conditioning is defined, we also have:


% \subsection{Further Axiomitization: Compatibility with Structure}
% \textit{DIVERGENCE.~~}
% \textbf{Divergence.}
\begin{wip}
\subsubsection*{Compatibility with Divergences and Cost Functions}
Suppose that, in addition, we have a ``divergence'' function $d : X \times X \to \mathbb R^+$ on $X$, with the property that $d(x,y) = 0$ iff $x = y$.
For sets $A \subset X$, let $d(x, A) := \inf_{a \in A} d(x,a)$ be the smallest possible divergence between $x$ and any member of $A$; symmetrically, define $d(A, x) := \inf_{a \in A} d(a,x)$.

% What we are trying to do with divergence functions here is an example of a
% slightly more broader definition,
If we drop the requirement that the smallest possible value of $d(A,x)$ must equal zero, we obtain the more general notion of a cost (or loss) function, $c :\mathcal A \times X \to \mathbb R_{\ge 0}$ satisfying
\[
% c : \mathcal A \times X \to \mathbb R_{\ge 0}
% \quad\text{such that}\quad
\forall A \in \mathcal A.~\arg\min_{x} c_A(x) = A.
% \forall A \in \mathcal A.\quad c_A(x) = 0 ~\Leftrightarrow~ x \in A.%
% \footnote{Any $c$ with the property that $\arg\min_{x} c_A(x) = A$ for all $A$ only differs from such a $c$ by a constant.}
% \qquad \text{and}\qquad
% \inf_{a \in A} c_A()
\]
The idea is that, if $c(x)$ is some cost that ``incentivises'' membership in $A$, then increasing confidence in $A$ ought decrease expected cost.
% Note that given a divergence function $d$ as described above, the expression $d(A, x) = \inf_{a \in A} d(a,x)$ satisfies this criterion.
\begin{CFaxioms}
    \item For all $\beta > 0$ and $A\in\mathcal A$, we have
    $\Ex_{F^{\beta}_A(\Pr)} [ c_A ]
        \le
        \Ex_{\Pr}[ c_A ]
    $
        \hfill \textbf{($c$\,-monotonicity)}
\end{CFaxioms}


We can also use $c$ to define a notion of independence.

\begin{defn}[$c$-independence]
For $A,B,Z \subset X$,
we say that $A$ and $B$ are $c$-independent given $Z$ iff
% $d(A, b) = d(a, B)$ for all $a\in A$ and $b \in B$.
for all $z \in Z$, we have that
% $c(z, A \cap B) = c(z, A) + c(z,B)$.
$c_{A\cap B}(z) = c_{A}(z) + c_{B}(z)$.
\end{defn}

To give a few examples:
\begin{enumerate}
    \item Every set $A \in\mathcal A$ is independent of the trivial event $X$, since
        $
            c(z, A\cap X) = c(z, A)
        $
        and $c(z, X) = 0$.
    \item Suppose $\X$ is a subset of $\mathbb R^n$, so that $x = (x_1, \ldots, x_n)$, and cost is L1 distance, given by $c(x,A) = \inf_{a \in A} \sum_{i=1}^n {|a_i - x_i|}$. Now for $i\ne j$,
    the sets $A_i(b) := \{ x : x_i = b \}$
    and $A_j(b') := \{x : x_j = b' \}$ are unconditionally $c$-independent.
    This makes sense, since they are orthogonal hyperplanes.

    % \item
    % \begin{wip}
    % Suppose $\X \cong Y \times Z$ is a product space, to be thought of as the set of possible joint settings of two variables $Y$ and $Z$.
    % Now, fix a probability measure $p(Y,Z)$ over $\X$
    % and let
    % % $c(x, A) := \I_{p|A}(\{x\}) = -\log P(x | A)$.
    % $c(x, A) := -\log p(A \cap \{x\})$, with the convention that $-\log 0 = \infty$.
    % Then, $A, B$ are $c$-independent given $Z$ iff
    % \[
    %     - \log p(A \cap B \cap \{z\}) =
    %     - \log p(A \cap \{z\}) - \log p(B \cap \{z\})
    % \]
    % If $z \notin A \cap B$, then both sides will be infinite. If $z \in A \cap B$,
    % \[
    %     - \log p(z) = - \log p(z) - \log p(z)
    %     \quad\iff\quad
    %     p(z) = p(z)^2
    %     \quad\iff\quad
    %     p(z) \in \{0,1\}
    % \]
    % \end{wip}
\end{enumerate}

% Once we have a notion of independence such as this one (although we will also consider others), we can articulate another axiom for update rules:
Armed with this notion of independence, we can articulate one further axiom for udpate rules:
\begin{CFaxioms}
% \begin{enumerate}[resume,label=\textbf{UR\arabic{*}.},nosep,leftmargin=2cm]
    % \item[\textbf{UR4}.]
    \item
    % \item[\textbf{UR4-d}.]
    % If $A$ and $B$ are $d$-independent, then
    % $F^{\beta}_A \circ F^{\beta}_B = F^{\beta}_{A \cap B}$.
    %     \hfill \textbf{($d$-decomposition)}
    If $A$ and $B$ are independent given $\mathrm{Supp}(\Pr)$, then
    $F^{\beta}_A \circ F^{\beta}_B (\Pr) = F^{\beta}_{A \cap B}(\Pr)$. \\
        \hfill \textbf{(decomposition)}
    % \item $\Ex_{F^\beta_A \Pr}[ d(X,A)]$ is (strictly) decreasing in $\beta$, so long as it is positive
    % \hfill \textbf{(monotonicity)}
% \end{enumerate}
\end{CFaxioms}
The idea here is that if $A$ and $B$ are independent, then it is equivalent to learn them in either order, or both at once.
\end{wip}
\begin{wip}

One possible choice of $\Phi$ statements with the $\sigma$-algebra $\mathcal A$ of measurable sets, which are closed under union and complementation, and on which probabilistic conditioning is defined, we also have:

\begin{CFaxioms}
    \item If $\Pr(A) > 0$ then $\displaystyle \lim_{\beta\to\infty} F^\beta_A (\Pr) = \Pr|A$
    \hfill \textbf{(absolute certainty)} \label{ax:certainty}


    \item
    % $F^\beta_A$ and $F_{\bar A}^\beta$ are inverses.
    $F^\beta_A \circ F_{\bar A}^\beta = 1_{\Delta\X}$.
        \hfill \textbf{(complementation)} \label{ax:comp}
\end{CFaxioms}


We can also consider the weaker variant of \Cref{ax:certainty}:
\begin{CFaxioms}
    % \item[U3$'$.]  \textbf{(effectiveness)~} $\supp F^\infty_A (\Pr) \subset A $
    % customlabel
    \item[\Cref*{ax:certainty}$'$]
    % \item[\customlabel{ax:effectiveness}{\textbf{\Cref*{ax:certainty}$'$.}}]
    If $\Pr(A) > 0$, then
        % $F^\infty_A (\Pr)(A) = 1$
        $\displaystyle \lim_{\beta\to\infty} F^\beta(\Pr)(A) = 1$
        % \makeatletter
        % \@currentlabel
        % \makeatother
        \hfill \textbf{(effectiveness)} \label[CFaxiomsi]{ax:effectiveness}
\end{CFaxioms}
in which the limit of infinite confidence in $A$ is not required be the same as conditioning the starting point on $A$, but still must have the same effect of ensuring that $A$ occurs with probability 1.



% \subsection{Further Axiomitization: Compatibility with Structure}
% \textit{DIVERGENCE.~~}
% \textbf{Divergence.}
\subsubsection*{Compatibility with Divergences and Cost Functions}
Suppose that, in addition, we have a ``divergence'' function $d : X \times X \to \mathbb R^+$ on $X$, with the property that $d(x,y) = 0$ iff $x = y$.
For sets $A \subset X$, let $d(x, A) := \inf_{a \in A} d(x,a)$ be the smallest possible divergence between $x$ and any member of $A$; symmetrically, define $d(A, x) := \inf_{a \in A} d(a,x)$.

% What we are trying to do with divergence functions here is an example of a
% slightly more broader definition,
If we drop the requirement that the smallest possible value of $d(A,x)$ must equal zero, we obtain the more general notion of a cost (or loss) function, $c :\mathcal A \times X \to \mathbb R_{\ge 0}$ satisfying
\[
% c : \mathcal A \times X \to \mathbb R_{\ge 0}
% \quad\text{such that}\quad
\forall A \in \mathcal A.~\arg\min_{x} c_A(x) = A.
% \forall A \in \mathcal A.\quad c_A(x) = 0 ~\Leftrightarrow~ x \in A.%
% \footnote{Any $c$ with the property that $\arg\min_{x} c_A(x) = A$ for all $A$ only differs from such a $c$ by a constant.}
% \qquad \text{and}\qquad
% \inf_{a \in A} c_A()
\]
The idea is that, if $c(x)$ is some cost that ``incentivises'' membership in $A$, then increasing confidence in $A$ ought decrease expected cost.
% Note that given a divergence function $d$ as described above, the expression $d(A, x) = \inf_{a \in A} d(a,x)$ satisfies this criterion.
\begin{CFaxioms}
    \item For all $\beta > 0$ and $A\in\mathcal A$, we have
    $\Ex_{F^{\beta}_A(\Pr)} [ c_A ]
        \le
        \Ex_{\Pr}[ c_A ]
    $
        \hfill \textbf{($c$\,-monotonicity)}
\end{CFaxioms}


We can also use $c$ to define a notion of independence.

\begin{defn}[$c$-independence]
For $A,B,Z \subset X$,
we say that $A$ and $B$ are $c$-independent given $Z$ iff
% $d(A, b) = d(a, B)$ for all $a\in A$ and $b \in B$.
for all $z \in Z$, we have that
% $c(z, A \cap B) = c(z, A) + c(z,B)$.
$c_{A\cap B}(z) = c_{A}(z) + c_{B}(z)$.
\end{defn}

To give a few examples:
\begin{enumerate}
    \item Every set $A \in\mathcal A$ is independent of the trivial event $X$, since
        $
            c(z, A\cap X) = c(z, A)
        $
        and $c(z, X) = 0$.
    \item Suppose $\X$ is a subset of $\mathbb R^n$, so that $x = (x_1, \ldots, x_n)$, and cost is L1 distance, given by $c(x,A) = \inf_{a \in A} \sum_{i=1}^n {|a_i - x_i|}$. Now for $i\ne j$,
    the sets $A_i(b) := \{ x : x_i = b \}$
    and $A_j(b') := \{x : x_j = b' \}$ are unconditionally $c$-independent.
    This makes sense, since they are orthogonal hyperplanes.

    % \item
    % \begin{wip}
    % Suppose $\X \cong Y \times Z$ is a product space, to be thought of as the set of possible joint settings of two variables $Y$ and $Z$.
    % Now, fix a probability measure $p(Y,Z)$ over $\X$
    % and let
    % % $c(x, A) := \I_{p|A}(\{x\}) = -\log P(x | A)$.
    % $c(x, A) := -\log p(A \cap \{x\})$, with the convention that $-\log 0 = \infty$.
    % Then, $A, B$ are $c$-independent given $Z$ iff
    % \[
    %     - \log p(A \cap B \cap \{z\}) =
    %     - \log p(A \cap \{z\}) - \log p(B \cap \{z\})
    % \]
    % If $z \notin A \cap B$, then both sides will be infinite. If $z \in A \cap B$,
    % \[
    %     - \log p(z) = - \log p(z) - \log p(z)
    %     \quad\iff\quad
    %     p(z) = p(z)^2
    %     \quad\iff\quad
    %     p(z) \in \{0,1\}
    % \]
    % \end{wip}
\end{enumerate}

% Once we have a notion of independence such as this one (although we will also consider others), we can articulate another axiom for update rules:
Armed with this notion of independence, we can articulate one further axiom for udpate rules:
\begin{CFaxioms}
% \begin{enumerate}[resume,label=\textbf{UR\arabic{*}.},nosep,leftmargin=2cm]
    % \item[\textbf{UR4}.]
    \item
    % \item[\textbf{UR4-d}.]
    % If $A$ and $B$ are $d$-independent, then
    % $F^{\beta}_A \circ F^{\beta}_B = F^{\beta}_{A \cap B}$.
    %     \hfill \textbf{($d$-decomposition)}
    If $A$ and $B$ are independent given $\supp(\Pr)$, then
    $F^{\beta}_A \circ F^{\beta}_B (\Pr) = F^{\beta}_{A \cap B}(\Pr)$. \\
        \hfill \textbf{(decomposition)}
    % \item $\Ex_{F^\beta_A \Pr}[ d(X,A)]$ is (strictly) decreasing in $\beta$, so long as it is positive
    % \hfill \textbf{(monotonicity)}
% \end{enumerate}
\end{CFaxioms}
The idea here is that if $A$ and $B$ are independent, then it is equivalent to learn them in either order, or both at once.


% \begin{enumerate}
%     \item[\textbf{UR4-c}.] If $A$ and $B$ are $d$-independent, then $F^{\beta}_A \circ F^{\beta}_B = F^{\beta}_{A \cap B}$.
%         \hfill \textbf{($c$-decomposition)}
% \end{enumerate}

%%%%%%% OTHER STRUCTURE
% \subsubsection*{A Topology}
% \subsubsection*{A Smooth Manifold Structure}
% \subsubsection*{A Matroid}
% \subsubsection*{A Riemannian Metric}
% If $\Delta X$

\subsubsection*{Following Geodesics}

There is a unique Riemannian metric $g$ on the space $\theta$ induced by the family $p_\theta$, which is invariant
under sufficient statistics \cite{}, called
the ``Fisher information metric''.
\[
    g_{ij} = .
\]

From this Riemannian metric, we obtain a natural notion of distance between probability distributions, and of a geodesic between them.

\begin{CFaxioms}
    \item Updates move along geodesics. That is, for any $\phi$, $\mu$, and $\beta$,
    the path $\gamma(t) = F^{t}_\phi(\mu)$ from $t = 0$ to $t = \beta$ has minimal arc length,
    among all paths from $\mu$ to $F^{\beta}_\phi(\mu)$.
\end{CFaxioms}



\subsubsection*{}

\begin{CFaxioms}
    \item If $\phi$ does not distinguish between events $A$ and $B$, then updating with $\phi$ does not change the relative probabilities of $A$ and $B$. That is, for all $\beta >0$,
    \begin{align*}
        \frac{ \Pr_{F^\beta_\phi \theta} (A) }{ \Pr_{F^\beta_\phi \theta} (B) }
        = \frac{ \Pr_{\theta} (A) }{ \Pr_{\theta} (B) }
            &&\textbf{(IIA)}
    \end{align*}

\end{CFaxioms}

But this may not always be reasonable. Consider the following example.
\begin{example}
    Consider the case of a binary outcome $X = \{0,1\}$,
    where $\Theta := [0,1]$ parameterizes all distributions, so that
    $\Pr_\theta(X=1) = \theta$, and the set of observations $\Phi := [0,1]$ indicates

    Now, if $\theta = 0.9$,
\end{example}


\end{wip}

\subsection{Elementary Properites of Update Rules}
\begin{prop}[closure under rescaling]
    If $F^{\beta}_\phi$ is an update rule for $\Phi$, then so is
    $G^{\beta}_\phi := F^{k \beta}_\phi$, for any positive real number $k > 0$.
        \label{prop:rescale}
\end{prop}


\textbf{Sequences.}
A $\Theta$-update rule $F$ on $\Phi$ also suggests an extension to updates on the more expressive set
\[
(\Phi \times \mathbb R)^* := \Big\{
    \text{finite sequences}~
    %     [(\phi_1, \beta_1), \ldots (\phi_n, \beta_n)]
    ~ (\phi_i, \beta_i)_{i \in [n]}
    ~\Big|~
    % \forall i \in \{1 \ldots n \}. ~ \phi_i \in \Phi, \beta_i \in \mathbb R
    ~ \phi_i \in \Phi, \beta_i \in \mathbb R \,\text{ for all }\, i \in
        % \mathbb N~
        [n]
    \Big\}
\]
via sequential composition of the underlying updates:
\[
    % \bar F^{\beta}_{\boldsymbol \varphi = (\beta_i \phi_i)_{i =1, \ldots n}}
    \bar F^{k}_{\boldsymbol \varphi \cdot \boldsymbol\beta}
        (\mu) := F^{k \beta_1}_{\phi_1} \circ F^{k \beta_2}_{\phi_2}
            \circ \cdots\circ F^{k \beta_n}_{\phi_n}(\mu).
\]

$\bar F$ satisfies \cref{ax:zero,ax:diffble,ax:symmetry} if $F$ does.
In general, it will not in general be an update rule, as it may not be additive (\cref{ax:additivity}). But if $F$ is commutative (\cref{ax:commute}), then $\bar F$ \emph{does} satisfy \cref{ax:additivity}, and is also commutative (\cref{ax:commute}) itself.
% , since
% \[
% \bar  F^{k'}_{\boldsymbol \varphi' \cdot \boldsymbol\beta'}
%     (\bar F^{k}_{\boldsymbol \varphi \cdot \boldsymbol\beta} (\mu))
%     =
%     F^{k' \beta'_1}_{\phi'_1} \circ F^{k' \beta'_2}_{\phi'_2}
%         \circ \cdots\circ F'^{k' \beta'_{n'}}_{\phi'_{n'}}
%     (
%         F^{k \beta_1}_{\phi_1} \circ F^{k \beta_2}_{\phi_2}
%         \circ \cdots\circ F^{k \beta_n}_{\phi_n}(\mu))
%     =
%         .
% \]


We will see that even non-commutative update rules, for which the order is important, can also be naturally combined in an unordered way.

\subsection{Vector Field Representations of Update Rules}
    \label{sec:vecrep}
For a smooth manifold $M$
(such as the space $\Delta \X$ of distributions over $\X$),
and a point $p \in M$, we follow convention by writing $T_p M$ for the tangent space to $M$ at point $p$ \parencite{lee2013smooth}, and % $TM := \sqcup_{p \in M} (p, T_p M)$
$TM := \sum_{p \in M} T_p M$ for the full tangent bundle over $M$.
%
A vector field over $M$ is a smooth map $\mat v : M \to T M$ assigning a tangent vector $\mat v(p) \in T_p M$, to every point $p \in M$.
% , or equivalently, a smooth section of the projection map $\pi : T M \to M$, where $\pi((p, v)) = p$.

\begin{theorem}\label{thm:vecrep}
    There is a natural bijection between
    % update rules $F : \Phi \times \mathbb R \to \Delta \X \to \Delta \X$
    update rules $F : \Phi \times \mathbb R \to (\Theta  \to \Theta)$
        satisfying \cref{ax:zero,ax:additivity,ax:diffble},
    and $\Phi$-indexed collections of vector fields
        % $\{ F'_\phi : \Delta X \to T \Delta X \}_{\phi \in \Phi}$%
        % $\{ F'_\phi : \Theta \to T \Theta \}_{\phi \in \Phi}$%
        $ F' :  \Phi \times \Theta \to T \Theta$%
        % $F' : \Phi \to \Delta\X \to T\Delta \X$%
    .
\end{theorem}

One defining feature of vector fields is closure under linear
    % (and, in particular, convex)
    combination.
Because they are in natural correspondance with differentiable additive update rules, update rules also inherit this structure.

In particular, given two update flows $F, G : \mathbb R \to \Theta$, we can define
$F \oplus G$ via the vector field $(F \oplus G)' = F' + G'$.

\TODO[TODO: prove impossible to individually they satisfy certainty, but not together.]

\begin{defn}
    For an assertion language $\Phi$, let $\ext\Phi$ denote
    the space of weighted formal sums of elements of $\Phi$.
    % the space $\mathbb R^{\Phi}_{\mathrm{fin}}$ of finitely supported vectors over $\Phi$,
\end{defn}

% If $\Phi$
\begin{prop}
    Every  update rule $F$ on $\Phi$, can be naturally extended to an update rule
    $\bar F$ on $\ext\Phi$
    % $\mathbb R^{\Phi}$,
    via the total vector field
    \[
        % \bar F'_{\vec{x}}( \theta ) := \sum_{\phi} F'_\phi(\theta) x_\phi.
        \bar F'_{\textstyle\sum_i a_i \phi_i} ( \theta ) := \sum_{i} a_i F'_{\phi_i}(\theta).
    \]
% \end{prop}
%
\end{prop}

If $\Phi$ is itself a measurable space, we can extend this further:
Every  update rule $F$ on $\Phi$, can be naturally extended to an update rule $\bar F$ on the space $\mathcal M(\Phi)$ of measures over $\Phi$, via
\[
    % \bar F'(\theta) := \sum_{} \beta_x \
    \bar F'_{\beta(\Phi)}( \theta ) := \int_{\Phi} F'_\phi(\theta) \mathrm d\beta.
\]




\begin{prop}[Update Rules as Dynamical System with Inputs]
    Differentiable update rules are naturally equivalent to control systems
    % $f : X \times Y \to TX$,
    which govern the behavior of functions
    \[
        \dot x(t) = f(x, u).
    \]
\end{prop}


Suppose you're trying to convince an agent of something, which is to say change their belief state so that it lies within some $S \subset \Theta$.
This now becomes a control problem, of supplying


% \subsection{}
\section{Various Properties Commmitment Functions Can Have}

\subsection{Commutative Update Rules}


All differentiable update rules are ``locally'' commutative, in the sense that the difference between
$F_{\phi_1}^\epsilon \circ F_{\phi_2}^\epsilon$ and
$F_{\phi_2}^\epsilon \circ F_{\phi_1}^\epsilon$ goes to zero as $\epsilon \to 0$.
This is an immediate consequence of differentiability and the fact that they share a limit point (the identity function).

If we fix a commutative and differentiable update rule $F$, and an initial point $\theta_0$, then the space $\mathbb R^\Phi$ of real-valued vectors over $\Phi$,
serves as a coordinate system for $\Theta$.


\begin{prop}
    In the setting where $\Theta = \Delta\X$
    % and $\Phi \subset \mathcal A$,
    cost-based decay update rules
    \[
        % F^{\beta}_A(\mu) \propto \mu(x) \exp(- \beta' c_{A}(x))
        F^{\beta}_\phi(\mu) \propto \mu(x) \exp(- \beta' c_{\phi}(x)),
    \]
    for a cost function $c : X \times \Phi \to \mathbb R$
    are commutative.
\end{prop}
\begin{lproof}
    For some normalization factors $Z, Z', Z''$, we have:
    \begin{align*}
         F^\beta_\phi( F^{\beta'}_{\phi'}(\mu))
         &= F^\beta_\phi \Big( \frac{1}{Z} \,\mu\, \exp(- \beta' c_{\phi'}) \Big) \\
         &= \frac{1}{Z'} \frac{1}{Z} \,\mu\, \exp(- \beta' c_{\phi'}) \exp(- \beta c_{\phi}) \\
         &= \frac{1}{Z''} \,\mu\, \exp(-\beta' c_{\phi'} - \beta c_\phi)
    \end{align*}
    which is the same expression when we exchange $(\phi, \beta)$ and $(\phi', \beta')$.
\end{lproof}

Note that this is true even for costs generated by asymmetric distances $c_{\{y\}}(x) = d(y, x) \ne d(x,y) = c_{\{x\}}(a)$.

Not all update rules of interest are commutative, even if otherwise well-behaved.

\begin{example}
    The inconsistency-reduction update rule, $\tau$, is not commutative, but it is differentiable, additive, invertable, and even conservative.
\end{example}


\subsection{Conservative Update Rules}

A \emph{conservative} vector field $Y : M \to T M$ on a differentiable manifold $M$ by be equivalently characterized as
\begin{itemize}
    \item ``path independent'', in the sense that, its line integrals for paths $\gamma : [0,1] \to M$, the line integral
    $\int_\gamma Y$ of $Y$ along $\gamma$
     depends only on the path endpoints $\gamma(0)$ and $\gamma(1)$.
    \item
        arising as the gradient of a potential function, i.e., $F = \nabla U$ for some $U : X \to \mathbb R$.
\end{itemize}
% Equivalently, conservative vector fields can be characterized as those fields so
The class of conservative vector fields is an incredibly important class.
It includes gravitational and electrostatic forces.

Recall that differentiable update rules are in 1-1 correspondence with indexed
collections of vector fields $\{ F'_\phi : \Theta \to T\Theta \}_{\phi \in \Phi}$,
% The corresponding update rules are therefore in correspondance with potential functions
\[
    U : \Phi \times \Theta\to \mathbb R
\]

% Any laws of mechanics, by Noether's theorem
% In three dimensions, they are called ``curl-free''.
% It is well-known that such vector fields are equivalently characterized as those
from which we get a vector field $F'_\phi(\theta) = \nabla_{\theta} U(\phi, \theta) \in T_{\theta} \Theta$.


\subsection{Linear Update Rules}

% In some sense, ALL update rules are linear in $\bar\Phi$ by definition.

There are many definitions of linear update rules:
\begin{defn}\label{ax:linear}
Let $F$ be a differentiable update rule on $\Theta$. We say that $F$ is \textellipsis
\begin{itemize}
    \item \emph{linear} if $\Theta$ is a vector space over $\mathbb R$, and the
    vector field $F'_\phi$ is a linear operator, i.e., for all $a, b \in \mathbb R$, we have that
    \[ F'_\phi(a \theta_1 + b \theta_2) = a F'_\phi(\theta_1) + b F'_\phi(\theta_2). \]

    \item \emph{cvx-linear} if $\Theta \subset \mathbb R^n$ is a convex set, and, for all $a \in [0,1]$, we have that
    \[ F'_\phi(a \theta_1 + (1-a) \theta_2) = a F'_\phi(\theta_1) + (1-a) F'_\phi(\theta_2). \]

    \item \emph{$\mathcal L$-cvx-linear} if $\Theta \subset \mathbb R^n$ and $F$ is an optimizing update rule with a loss representation $\mathcal L$ linear in its first argument, i.e.,
    \[
        \mathcal L(a \theta_1 + (1-a) \theta_2, \varphi) = a \mathcal L(\theta_1, \varphi) + (1-a) \mathcal L(\theta_2, \theta).
    \]
\end{itemize}
    % $F'_\phi(\theta) = \mathrm{V}_\phi \theta$ for some linear operator $V_\phi \in \mathbb R^{n \times n}$.
    % $F'_\phi(\theta) = \mathrm{V}_\phi \theta$ for some linear operator $V_\phi$.
\end{defn}

\begin{prop}
    If $F$ is a $\mathcal L$-cvx-linear, then it is also cvx-linear.
\end{prop}

In fact, the first condition is much stronger;
\begin{prop}
    if $F$ is a nontrivial $\mathcal L$-cvx-linear optimizing UR, then $\Theta$ equals cone generated by  the rays $\{ F'_\varphi\theta : \theta \in \Theta, \varphi \in \Phi \}$. In particular, if there is some $\theta$ such that $0$ is in the interior of the convex hull $\mathrm{conv}(\{F'_\phi\theta\}_{\phi \in \Phi})$, then $\Theta = \mathbb R^n$.
\end{prop}

% Implicit in this definition is the supposition that the integral curves generated by the differential equations, started at any point $\theta \in \Theta$, are

\begin{prop}
    % If $F$ is a  differ
    Every linear update rule is of the form
    $
        F^{\beta}_\phi(\theta) =  \theta^{T} \exp(\beta V)
    $,
    where $\exp(\beta V)$ is the matrix exponential.%
        \footnote{Concretely, if $V = U^T \mathrm{Diag}(\lambda_1, \ldots \lambda_n) U$ is an eigendecomposition of $V$, then $\exp(V) = U^T \mathrm{Diag}(e^{\beta\lambda_1}, \ldots e^{\beta\lambda_n}) U$.}
\end{prop}

\begin{prop}
    A linear update rule $F$ is commutative iff, for every pair of statements  $\phi, \phi' \in \Phi$, the
    matrices $V_\phi$ and $V_{\phi'}$ commute.
\end{prop}



\subsection{Invertable Update Rules}
\begin{CFaxioms}
    \item For all $\phi\in\Phi$, and $\beta \in \mathbb R$, the update
    $F^{\beta}_{\phi}: \Theta \to \Theta$ is invertable.
    \hfill\textbf{(Invertability)} \label{ax:invert}
\end{CFaxioms}


This effectively partitions $\Theta$ into two


\begin{prop}
    If $F$ is a differentiable and invertable update rule (i.e., satisfies \cref{ax:zero,ax:additivity,ax:invert,ax:diffble}), then for all $\beta \in \mathbb R$, $\phi \in \Phi$, the function
    % $F^\beta_\phi : \Delta\X \to \Delta\X$
    $F^\beta_\phi : \Theta \to \Theta$
    is a diffeomorphism, and its inverse is given by $F^{-\beta}_\phi$, in the sense that
    \[
        F^{-\beta}_\phi( F^{\beta}_\phi (\mu) ) = \mu = F^{\beta}_\phi( F^{-\beta}_\phi (\mu) ).
     \]
\end{prop}


% Together with strong additivity, we get
% \begin{prop}
%     If $F$ is an update rule satisfying \cref{ax:additivity,ax:invert},
%     then any update prescribed by $F$ (or sequence thereof) takes positive distributions to positive distributions.
%     %
%     Concretely, for all $\beta$ and $\phi$,   $\mu \in \mathrm{Int}(\Delta\X)$ if and only if $F^{\beta}_\phi(\mu) \in \mathrm{Int}(\Delta\X)$.
%     If $F$ further satisfies \cref{ax:sufficiency, ax:diffble}, then
%     % \[
%         $F^\beta_\phi$ is a diffemorphism of $\mathrm{Int}(\Delta\X)$.
%     % \]
% \end{prop}


As a consequence,
\begin{coro}
    If for any $\beta < \infty$ there exist $\mu, \phi, A$ such that
    $\mu(A) > 0$  but $F^{\beta}_\phi(\mu)(A) = 0$, then $F$ is not invertable.
\end{coro}

% \begin{CFaxioms}
%     \item
%     % The distributions equals the set of
%     \[
%         F^{\boldsymbol \beta} = \# \{ \phi : \beta_\phi = \infty \}
%     \]
%
% \end{CFaxioms}

% \section{A Canonical Update Rule}
\section{Optimizing Update Rules}
Suppose we have
\begin{enumerate}[nosep]
    \item A differentiable function $U : \Theta \times \Phi  \to \mathbb R$, which intuitively measures ``incompatibility'' between a belief state $\theta$ and an assertion $\varphi$, and
    \item
        % A way of taking gradients of $U$ with respect to $\theta$,
        % % $\nabla : ()$
        % such as an inner product $g_p : T_p\Theta \times T_p\Theta \to \mathbb R$, making $(\Theta, g)$ a Riemannian manifold.
        A way of taking the gradient of $U$ with respect to $\theta$.%
            \footnote{
            such as a tangent-cotangent isomorphism $(-)^\sharp : T^*_p\Theta \to T_p \Theta$, perhaps coming from an affine connection, in turn perhaps coming from a Riemmannian metric.}
\end{enumerate}
% \def\GD#1{(\mathtt{GD}\;#1)}
\def\GD#1{\mathtt{GF}[#1]}
\def\NGD#1{\mathtt{NGF}[#1]}

Then we can define an update rule $\GD U$ that reduces inconsistency by gradient descent. Concretely, such an update rule has a vector field:
% \def\GD#1{(\mathtt{GD}\; #1)}
\[
    \GD U'_\phi(\theta) = - \nabla_\theta U(\theta,\phi).
\]


\begin{prop}
    An update rule $F$ on a Riemannian manifold $\Theta$ is optimizing update rule if and only if $(F')^\flat$ is a conservative co-vector field.
    \cite[Prop 11.40]{lee2013smooth}
\end{prop}

\textbf{Natural Gradients for Probability Distributions.}
When $\Theta$ parameterizes a family of probability distributions, via some $\Pr : \Theta \to \Delta \X$, there is a particularly natural metric on $\Theta$, called the Fisher information metric.
This metric is the unique one on $\Theta$ that is
independent of the representation of $\X$ \parencite{chentsov}, in the following sense.
If there are cpds $p(Y|X)$ and $q(X|Y)$ such that, for all $\theta \in \Theta$,
% $\Pr_\theta = qp\Pr(\theta)$,
% for all $\theta$, sampling $x\sim \Pr_\theta$
% is the same as the distribution over $x'$  $y \sim p(Y|x)$, and $x'\sim q(X|y)$, is equialent to samkp
the distribution $\Pr_{\theta}(X)$ is unchanged after converting to $Y$ and back again $X$ (via $p$ and $q$ respectively), as depicted by the following commutative diagram,
% $q \circ p \circ \Pr_\theta(X) = \Pr_\theta(X)$
\[
    % g [ \Theta \xrightarrow{\Pr} X ]  = g [ \Theta \xrightarrow{\Pr} X \xrightarrow{p}
    %  \Theta \xrightarrow{\Pr} X
    % \quad = \quad {} \xrightarrow{\theta} \Theta \xrightarrow{\Pr} X \overset p\to Y \overset q\to X,
    \begin{tikzcd}
        \Theta \ar[r,"\Pr"]\ar[d,"\Pr"']
            % \ar[rd,dashed,"\Pr^{(Y)}"description]
            & X \\
        X \ar[r,"p"'] & Y \ar[u, "q"']
    \end{tikzcd}
\]
then clearly the family $\Pr(Y|\Theta) := p\,\circ\,\Pr_{\theta}$ carries the same information about the parameters (and in particular how best to update them) as $\Pr_\theta$.
% since we can (losslessly) convert between the two.
Chentsov's theorem says, that, up to a multaplicative constant, the Fisher information metric is the only metric on $\Theta$, as a function of the parameterization $\Pr$, which gives identical geometry in both cases.
% $g[\Pr] = g[\Pr^{Y}]$
% \footnote{at least when $X$ and $Y$ take values in a finite set, although there have since been numerous extensions of it.}


% This allows us to compute the gradient with respect to this metric as
At each point $\Theta$, the components of the Riemannian metric form a matrix---in this case, the Fisher information matrix $\mathcal I(\theta)$---which allow us to now compute the gradient in the natural geometry from the coordinate derivatives as
\[
    \NGD {\cal L}'_\phi (\theta) = - \hat \nabla_\theta \mathcal L(\theta,\phi)
        % = \mathcal I(\theta)^{-1} \frac{\partial}{\partial \theta_i} U(\theta, \phi)
        = \mathcal I(\theta)^{\dagger}  \nabla \mathcal L(\theta, \phi)
\]
where $ \mathcal I(\theta)^{\dagger} $ denotes the Moore-Penrose psuedoinverse of the matrix $ \mathcal I(\theta)$,
and $\nabla \mathcal L$ is the gradient for the euclidean metric i.e., the vector of partials $[\frac{\partial \mathcal L}{\partial \theta_1}, % \ldots, \frac{\partial U}{\partial \theta_i},
 \ldots, \frac{\partial \mathcal L}{\partial \theta_n}]^{\mathsf T}$.

\TODO


\subsection{Boltzmann Update Rules}
% Again, suppose we have a differentiable function $U : \Theta \times \Phi  \to \mathbb R$.
% Suppose further that we have a recovering a parameter that gives rise to a given probability distribution, i.e., a section of $\Pr$, or concretely,
% a function $\Pr^{-1} : \Delta\X \to \Theta$ such that $\Pr(\Pr^{-1}(\mu)) = \mu$ for all $\mu \in \Delta\X$.
%
% This time, define an update rule directly, by
%
% \begin{align*}
%     (\mathtt{Boltz} U)_\varphi^\beta(\theta)
%         &:\propto \Pr\nolimits_\theta \exp(-\beta U(\theta,\varphi))\\
%         &:= \Pr\nolimits^{-1}\bigg\{
%             A \mapsto \Pr\nolimits_\theta(A)  \, \frac
%             % {1}
%             { \exp(-\beta U(\theta,\varphi)) }
%             {\Ex_{\Pr_\theta}[\exp(-\beta U(\theta,\varphi))]}
%             % \int_A  \exp(-\beta U(\theta,\varphi)) \mathrm d\,\Pr\nolimits_\theta
%         \bigg\}
% \end{align*}
% Now suppose we have a potential function $U : X \times \Phi  \to \mathbb R$.
\def\Bolz#1{\mathrm{Bolz}[#1]}
Suppose, for each $\phi \in \Phi$, we have a potential function $U_\phi : X \to \mathbb R$ on the underlying set $X$.
We can use this to define an update rule
\begin{align*}
    \Bolz U &: (\mathbb R \times \Phi) \to \Delta\X \to \Delta\X \\
    \Bolz U^\beta_\varphi(\mu)
        &:\propto
            \mu \exp(-\beta U_\phi) \\
        &= A \mapsto \frac
            1{\Ex_{\mu} [ \exp(-\beta U_\phi)]}
            {\int \exp(-\beta U_\phi) \mathbbm 1_A \mathrm d\mu }
\end{align*}

\begin{prop}
    Bolzmann Update Rules are additive, zero, differentiable, invertable, and commutative.
\end{prop}

\begin{lproof}
    \textbf{Commutativity.}
    For some normalization factors $Z, Z', Z''$, we have:
    \begin{align*}
         F^\beta_\phi( F^{\beta'}_{\phi'}(\mu))
         &= F^\beta_\phi \Big( \frac{1}{Z} \,\mu\, \exp(- \beta' c_{\phi'}) \Big) \\
         &= \frac{1}{Z'} \frac{1}{Z} \,\mu\, \exp(- \beta' c_{\phi'}) \exp(- \beta c_{\phi}) \\
         &= \frac{1}{Z''} \,\mu\, \exp(-\beta' c_{\phi'} - \beta c_\phi)
    \end{align*}
    which is the same expression when we exchange $(\phi, \beta)$ and $(\phi', \beta')$.
\end{lproof}

Note that this is true even for costs generated by asymmetric distances $c_{\{y\}}(x) = d(y, x) \ne d(x,y) = c_{\{x\}}(a)$.


\begin{remark}
    % For fixed, $\varphi$,
     % and in particular, if $\Phi = \{\varphi\}$ is a singleton,
    Regarding $U_\varphi : \X \to \mathbb R$ as a potential energy over $X$,
    $\Bolz U^\beta_\varphi(\Unif)$ is the Boltzmann distribution at inverse temperature (thermodynamic coldness) $\beta$
    % and base measure $\mu$.
    In the thermodynamic analogy, as temperature decreases, one becomes more certain that particles are in their most favorable states.
\end{remark}


The certainties of $\Bolz U$ are the minimizers of $U$.


\begin{wip}
    % Suppose $\Theta$
    As a reminder, we have $\Theta = \Delta \X$, and suppose we have $c : X \times \Phi \to \mathbb R$.
    Suppose $U(\theta, \varphi) = \Ex_\theta[ c_\varphi ]$ (linearity, \cref{ax:linear}).
    Then
    $% \begin{align*}
        (\mathrm{Boltz}\;U)'_\varphi \theta = \theta(\Ex_\theta[c_\varphi] - c_\varphi),
    $% \end{align*}
    while
    \begin{align*}
         (\mathtt{GD}\; U)'_\varphi \theta &=
             - \nabla_{\theta} \Ex\nolimits_\theta [ c_\varphi ]
             \\ &= - c_\varphi
             % {\color{gray}~+ \Ex\nolimits_{\theta}[c_\varphi]}
             .
    \end{align*}
    This second expression, though, doesn't seem quite right --- it isn't even a tangent vector to the probability simplex, since its components don't sum to zero.
    This issue is in our naive computation of the gradient.
    We have computed the collection of partial derivatives $\frac{\partial}{\partial \theta_i} ( \theta \cdot  c_\varphi) = (c_\varphi)_i$,
    which is technically  co-vector field, not a vector field.%
        \footnote{it acts on a vector field $V$ by $ -c_\varphi ( V ) = - V(c_\varphi)$.}

    For simplicity, suppose that $X = \{1, \ldots, n\}$, in the discussion that follows.
    If our parameter space were all of $\mathbb R^n$, we could simply collect these terms and take a transpose, to get
    \[
        \nabla_\theta \Ex_{\theta}[c_\varphi]
    \]
    There are two ways to proceed from here.
    The first makes use of manifold theory: for each point $p \in \Delta X$, begin by identifing a neighborhood $U \ni p$ with an open subset of $\mathbb R^{(n-1)}$, and define an inner product (a metric tensor) $g_p(\cdot, \cdot)$ on tangent vectors $v \in T_p\Delta X$, making $\Delta\X$ into a Riemannian Manifold, and then compute the gradient in the standard way, using the inverse of the metric tensor $g$ in order to convert covectors to vectors in a natural way.

    % The second, which is simpler but more constrained than the first, is to simply
    The second, which is computationally simpler, is to take the metric induced by an embedding in Euclidean space.
    This approach is equally general, because Nash's Theorem \parencite{nash1956imbedding} tells us that any $n$-dimensional Riemannian manifold may be isometrically embeded in $\mathbb R^{2n+1}$.



    % we have
    % \[ \frac{\partial}{\partial \theta^i} ( \theta \cdot  c_\varphi) \]
\end{wip}

\begin{prop}
    % Fix $\varphi$.
    % Let $f(X) := \exp(-\beta U(X,\varphi))$, and $g(X) := U(X,\varphi)$.
    % With the metric on $\Delta\X$ induced by its embedding as a simplex in $\mathbb R^{|X|}$, we have that
    % When parameterized as a simplex,
    The associated vector field is given by
    %
    % \begin{align*}
    $
        % (\mathrm{Boltz}\,U'_\varphi p)_{x} = p(x) (\Ex_p[U(X,\varphi)] - U(x,\varphi) )
        (\mathrm{Boltz}\,U)'_\varphi p = p (\Ex_p[U_\varphi] - U_\varphi )
    $.
    % \end{align*}
\end{prop}
\begin{lproof}
    Let $f(X) := \exp(-\beta U(X,\varphi))$, and $g(X) := U(X,\varphi)$.
    \begin{align*}
        \mathrm{Boltz}'_\varphi\theta &= \frac{\partial}{\partial \beta} \mathrm{Boltz}^\beta_\varphi(p) \Big|_{\beta=0} \\
    \intertext{\TODO[TODO: finish typesetting algebra]}
        &= x \mapsto
            p(x) \frac{f(x)}{\Ex_p[f]}
                \left(\Ex_p\left[ \frac{f}{\Ex_{p}[f]} g\right] - g(x) \right)
                % {\exp(-\beta U(x,\varphi))}
                % {\Ex_{}}
                \Big|_{\beta=0}
                \\
        &= \frac{pf}{\Ex_p[f]^2}
            \left(\Ex\nolimits_p\left[ f g\right] - g \Ex\nolimits_{p}[f] \right)
            \Big|_{\beta=0} \\
        &= x \mapsto p(x) (\Ex\nolimits_p[g] - g(x)) &
            \text{since $f(X) = 1$ when $\beta=0$}
    \end{align*}
    As a sanity check, note that the sum over all components is
    \[ \sum_{x \in X} ((\mathrm{Boltz}\,U)'_\varphi\, \theta)_x
         = \sum_{x \in X} p(x) (\Ex\nolimits_p[g] - g(x))
         = \Ex\nolimits_p[ \Ex\nolimits_p [ g ]] - \Ex\nolimits_p [g] = 0,
     \]
     so indeed it lies within the tangent space.
\end{lproof}


\begin{prop}
    The Bolzmann update rule with potential $U(X)$ is the natural gradient flow update rule for expected value of $U$, i.e.,
    \(
        \Bolz U =
        \NGD{\mu\mapsto\Ex_\mu U} .
    \)
\end{prop}




% \newpage%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
\subsection{Assorted}

\begin{examplex}{Weighted Average}{g1}
    \begin{align*}
        \Theta = \mathbb R^n \times (\mathbb R_{> 0} \cup \{\infty\});
        \qquad
        \Phi = \mathbb R^n %; \qquad
    \end{align*}
    A belief state $(\mat x,w) \in \Theta$ consists a current estimate $\mat x$ of the quantity of interest, and a weight $w$ of the total internal confidence in the estimate.

    Updating proceeds by taking a weighted average of the previous estimate and the new input, weighted by their respective confidences, which is captured by:
    \begin{align*}
        F^\beta_{\mat y}(\mat x, w) &=  \left( \frac{ w \mat x + \beta \mat y}{w + \beta} , w + \beta \right)
        \quad\text{and}\quad
        F^{\beta}_{\mat y}(\mat x, \infty) = (\mat x, \infty)
    \end{align*}
    It is additive, since
    \begin{align*}
        F^{\beta_2}_{\mat y} \circ F^{\beta_1}_{\mat y}(\mat x, w)
        &= \left( \frac{ (w + \beta_1) \frac{ w \mat x + \beta_1 \mat y}{w + \beta_1} + \beta_2 \mat y}{(w + \beta_1) + \beta_2}, (w  + \beta_1) + \beta_2 \right) \\
        &= \left( \frac{  w \mat x + (\beta_1 + \beta_2) \mat y}{w + (\beta_1 + \beta_2)}, w  + (\beta_1 + \beta_2)\right)
        = F^{\beta_1 + \beta_2}_{\mat y}(\mat x, w)
    \end{align*}
    And it is clearly differentiable, with a simple calculation revealing that
    $ F'_{\mat y}(\mat x, w) = \left( \frac{\mat y - \mat x}{w}, 1\right) $.

    Observations:
    \begin{itemize}
        \item The update rule cannot be extended differentiably to states $\theta = (\mat x, w)$ with $w = 0$.
            Intuitively, we need to have some estimate with positive confidence to update beliefs in a differentiable way.
            This is related to the fact that plain emperical risk minimization (ERM) is unstable, but stable with even a small amount of regularization.
            % It is also similar to the fact that one needs a prior in order to do
            % In this case, the prior may be arbitrarily weak.
        \item The certainties are given by
        \[
            \lim_{\beta \to \infty} F^{\beta}_{\mat y}(\mat x, w) = (\mat y, \infty)
        \]
        % \item In this case,
        \item $F$ is commutative, invertible, and symmetric with respect to permutation of the dimensions, but it is not conservative: if we had $U(\mat x, w, \mat y)$ twice differentiable such that $\nabla_{\mat x, w} U = F'$, then we would have
        \[\frac{\partial^2}{\partial w \partial x_i} U =
            \frac\partial{\partial w} \frac{y_i - x_i}{w} = \frac{x_i - y_i}{w^2}
            \qquad
            \text{but}
            \quad
            \frac{\partial^2}{\partial x_1 \partial w} U
            = \frac\partial{\partial x_i} 1 = 0
        \]
        violating Clairaut's theorem on equality of mixed partials.
        Therefore, $F$ is not an optimizing update rule.
    \end{itemize}
\end{examplex}
\begin{examplex}{Sensor Fusion, v1}{sfv1}
    Suppose we have a sensors, which measures a variable $X$, and gives an estimate $S$.
    Suppose that we also have a prior probability $q(X|S)$ of the value of the variable given the value of the sensor, and also a
\end{examplex}
\begin{examplex}{Sensor Fusion, Many Variables}{sfv2}
    Suppose we have $n$ sensors, which perorm measurements $\{S_1, \ldots S_n\}$ of variables $\{X_1, \ldots, X_n\}$, but may not be entirely accurate.
\end{examplex}

\begin{examplex}{Gaussian NGD}{gauss-ngd}
    Consider the case where $\Theta  = \{ (\mu, \sigma^2) \in \mathbb R \times \mathbb R_+ \}$ is the half-space of parameters to a Gaussian over some real variable $X$, and $\Phi \cong \mathbb R$ consists of possible observations of $X$.

    One natural loss function is negative log likelihood (differential surprisal) of the observation $x$ according to your belief state $\theta = (\mu, \sigma^2)$:
    \[
        \mathcal L(x, \mu, \sigma^2) = - \log \mathcal N(x \mid \mu, \sigma^2) = \frac12 \log 2\pi\sigma^2  + \frac12 \left(\frac{x-\mu}{\sigma}\right)^2 =
        \aar*{%
        \begin{tikzpicture}[center base]
            \node[dpad0](X) at (0,0){$X$};
            % \node[dpad
            \draw[arr2, <<-] (X) to node[above,pos=0.7]{$x$} ++(1.0,0);
            % \draw[arr2, <-] (X) to node[above]{$\mathcal N(X|\mu,\sigma^2)$} ++(-1.7,0);
            \node[dpad0](m) at (-1.5, 0.3) {$\mu$};
            \node[dpad0](s2) at (-1.5, -0.3) {$\sigma^2$};
            \mergearr{m}{s2}{X};
            \node[above=2pt of center-ms2X](N){$\mathcal N$};
            \draw[arr1, <<-] (m) to node[above, pos=0.7]{$\mu$} +(-0.9,0);
            \draw[arr1, <<-] (s2) to node[above, pos=0.7]{$\sigma^2$} +(-0.9,0);
        \end{tikzpicture}}.
    \]

    The fisher information for a normal distribution is given by
    \[
    \mathcal I(\mu, \sigma) =
    % \begin{bmatrix}
    %     \Ex_{\mathcal N(X|\mu,\sigma^2}[ \frac{\partial mathcal N(X|\mu,\sigma^2}{\partial \mu}) ]
    % \end{bmatrix}
    % =
    \begin{bmatrix}
        \frac1{\sigma^2} & 0 \\
        0 & \frac{1}{2 \sigma^4}
    \end{bmatrix}
    \]



    The natural gradient update rule is given by
    \[
        F'_{x}(\mu, \sigma^2) = - \hat\nabla_{\mu, \sigma^2} \mathcal L(x,\mu,\sigma^2) = \mathcal I(\mu, \sigma^2)^{-1}
        \begin{bmatrix}
            \frac{x-\mu}{\sigma} \\[1ex] \frac {-\sigma^2 + (x-\mu)^2}{2 \sigma^4}
        \end{bmatrix}
        =
        \begin{bmatrix}
            x-\mu \\ (x-\mu)^2 - \sigma^2
        \end{bmatrix}
    \]

    Note that:
    \begin{itemize}
        \item  $\Ex_{x \sim \nu} [ F'_x(\mu,\sigma^2) ] = \mat 0$ if and only if $\nu$ has mean $\mu$ and variance $\sigma^2$.
        % We can interpret this in two ways:
        % This means that if observations are drawn from a fixed distribution $\nu(X)$,
        Moreover, this point is the unique global attractor.
        This means that,
        \begin{enumerate}
            \item If observations are drawn from a fixed distribution $\nu(X)$, and we repeatedly use $F$ to update $\theta = (\nu, \sigma)$ with small confidence $\epsilon$,
            then $\mu$ will approach the mean $\Ex_{\nu}[X]$ of $\nu$
            and $\sigma^2$ will approach the variance $\Ex_{\nu}[X^2] - \Ex_{\nu}[X]^2$.
            % If we make f observations are drawn from a fixed distribution $\nu(X)$,
            % If we update our belief parameters $\theta = (\mu, \sigma)$ according to $F$ with

            \item If we perform a single high-confidence update on the extended observation $\varphi \propto \nu$, in which each $x$ has relative confidence $\nu(x)$, the result will be a Gaussian with the mean and variance of $\nu$, i.e.,
            \[
                \forall \theta.\quad
                \lim_{c \to \infty} \Pr\nolimits_{F_{\nu}^c(\theta)} = \mathcal N(\Ex\nolimits_{\nu}[X], {\mathrm{Var}}_{\nu}[X])
            \]
        \end{enumerate}
        In this sense, relative confidence acts like probability.
        % So,

        \item
        If we update with the observation $x = \mu$ of our estimate with confidence $c$,
        the mean is unchanged, and our estimate of the variance becomes the harmonic mean of our previous variance $\sigma_0^2$ and the inverse confidence $\frac1c$.
        That is,
        \[
            F^c_\mu(\mu, \sigma_0^2) =
            \left(\mu, \frac{1}{c + \frac{1}{\sigma_0^2}} \right).
        \]
        In particular, if $\sigma_0^2$ is very large, so that our initial beliefs say very little, updating with confidence $c$ results in variance $\frac1t$.
        In this sense, the magnitude of confidence acts as the inverse of variance.
    \end{itemize}
\end{examplex}


\begin{examplex}{Quantum}{q}
    % Let $W = \{1, \cdots, n\}$ be a finite set, and let 
    Let
    $\Theta$ be the set of possible quantum states over $n$ states, which 
    can be described as density matrices
    \[ 
        \Theta := \Big\{ \text{positive semi-definite matrices } \theta \in \mathbb R^{n \times n} \text{ such that } \Tr(A) = 1 \Big\}.
    \]

    Suppose also that 
    \[
        \Phi := \Big\{ \text{Hermitian Matrices } H \in \mathbb C^{n \times n} \Big\}
    \]

    The natural update rule here is 
\end{examplex}


\begin{prop}[Update Rules as Dynamical System with Inputs]
    Differentiable update rules are naturally equivalent to control systems
    % $f : X \times Y \to TX$,
    which govern the behavior of functions
    \[
        \dot x(t) = f(x, u).
    \]
\end{prop}

\subsection{Bayes Rule: Updating by Conditioning}
Fix some measurable space $\X = (X, \mathcal A)$ of possible outcomes, and consider an update rule for probability distributions $\Theta := \Delta\X$
on measurable events $\Phi = \mathcal A$.

The function
\[
    \mathrm{Bayes}^\beta_A(\mu) := \begin{cases}
            \mu \mid A &  \text{if }\beta > 0 \\
            \mu & \text{if } \beta = 0 \\
            \mu \mid \bar A &  \text{if } \beta < 0
        \end{cases}
\]
satisfies \cref{ax:zero,ax:additivity}, so it is an update rule.  But it is not differentiable in $\beta$ (it doesn't satisfy \cref{ax:diffble}).
% owever, it can be written as a limit of one that does:
% However, this behavior does arise as the limit of large $\beta$ for a differentiable update rule:
However, this behavior does arise as the limit of large-magnitude $\beta$ for a differentiable update rule. Consider the update rule
\begin{align*}
    F^\beta_A (\mu) &:\propto \mu \exp(\beta \mathbbm1[A]) \\
    F^\beta_A (\mu)(B) &:=
        % \mu(B) \exp(\beta \mathbbm1[A \cap B])
        \frac{1}{\Ex_{\mu}[e^{\beta\mathbbm1[A]}]}\int_{X}  \mathbbm1[B] \exp(\beta \mathbbm1[A]) \mathrm d\mu \\
        &=
        \frac{\mu(B\cap A)e^{\beta} + \mu(B \cap \bar A) }
        % {\mu(A)e^{\beta} + (1 - \mu(A))}
        {\mu(A)e^{\beta} + \mu(\bar A)}.
        % \int_{X}  \mathbbm1[B] \exp(\beta \mathbbm1[A]) \mathrm d\mu
\end{align*}

% It is easy to check that $F$ satisfies
% \cref{ax:zero,ax:additivity,ax:certainty,ax:symmetry,ax:commute,ax:diffble,ax:invert}.
% In addition to being a flow, this update rule
Note that
\begin{align*}
\lim_{\beta\to\infty} F^\beta_A(\mu) &= B \mapsto \frac{\mu(B \cap A)}{\mu(A)}
   = \mu \mid A; \\
 F^0_A(\mu) &= B \mapsto \mu(B); \qquad\text{and}\\
\lim_{\beta\to-\infty} F^\beta_A(\mu) &= B \mapsto \frac{\mu(B \cap \bar A)}{\mu(\bar A)}
      = \mu \mid \bar A.
\end{align*}
Thus, $\mathrm{Bayes} = \lim_{k \to \infty} F^{k}$, where $F^k$ is the update rule given by rescaling, as in \cref{prop:rescale}.



\begin{wip}
Is this the only such relaxation of Bayes Rule?

Suppose $F$ is another such relaxation, satisfying \cref{ax:zero,ax:additivity,ax:certainty,ax:symmetry,ax:commute,ax:diffble,ax:invert}.

\begin{prop}
    $F$ is unique up to curve reparameterization.
\end{prop}

By differentiability (\cref{ax:diffble}) and \cref{thm:vecrep}, we know that $F$ is generated by vector field $F' : \mathcal A \times \Delta\X \to \Delta\X$.
We start by investigating the action on singletons.
\[
    F'_{\{x\}}(\mu)
\]

Assuming that the vector field is conservative,

\begin{conj}
    $F$ is the unique update rule, up to multiplicative scaling, satisfying
    \cref{ax:zero,ax:additivity,ax:certainty,ax:symmetry,ax:commute,ax:diffble,ax:invert}
\end{conj}
\end{wip}




\subsection{Opacity and Image Blending}
Fix positive integers $W$ and $H$, and
let $\Theta := [0,1]^{W\times H\times 3}$ be the set of $W\times H$ colored images.
%
% Let $\Phi$
We'll consider several different sets of observations
\begin{align*}
     \Phi_0 &:= \Big\{ \text{Full Images}\quad \mat X \in [0,1]^{W \times H \times 3} \Big\} \\
     \Phi_1 &:= \Big\{ \text{Pixels at Given Positions}\quad [0,1]^3 \times (W \times H) \Big\} \\
     \Phi_1 &:= \Big\{ \text{Pixels at Given Positions}\quad [0,1]^3 \times (W \times H) \Big\} \\
\end{align*}

% First, exponential blending
In each case, we'll consider an ``expontial'' blending rule:
\[
    F^{\beta}_{\mat X}(\theta) :=
\]



\subsection{The Two-Point Space}
First, suppose $X = \{0,1\}$ and $\mathcal A = 2^X$.

\begin{claim}
The space of update rules satisfying
% UR1-6
\cref{ax:zero,ax:additivity,ax:certainty,ax:symmetry,ax:diffble}
% is in bijection with continuous monotone functions
is in bijection with the space of strictly increasing differentiable functions
$\gamma : [0,\infty) \to [0,1]$
% in the space $\Delta X \cong [0,1]$ of probability distributions over $\{0,1\}$,
such that
% \begin{align*}
%     \gamma_0(0) &= 1
%     &
%         \gamma_1(0) &= 0
%         \\
%     \lim_{t\to\infty}\gamma_0(t) &= 0
%     &
%         \lim_{t\to\infty}\gamma_1(t) &= 1
%         \\
% \end{align*}
$\gamma(0) = 0$ and $\lim_{t\to\infty}\gamma(t) = 1$.
\end{claim}
\begin{lproof}
    Let's first investigate the behavior of an update rule $F$ on the singleton $A := \{1\}$.
    % Fix some $\epsilon > 0$.
    % By \Cref{ax:zero,ax:additivity}
    % \ref{ax:effectiveness}
    % \cref{ax:effectiveness}
    By \ref{ax:effectiveness} or \cref{ax:certainty} we know that, for any $p \in \Delta\{0,1\}$, and particularly for $p$ equal to the degenerate distribution $\delta_0$ that places all mass on the outcome $X=0$, we have
    \[
        \lim_{\beta \to \infty} F^\beta_{\{1\}}(\delta_0) = \delta_1.
    \]
    Define $\gamma(t) := F^{t}_{\{1\}}(\delta_0)(X=1)$.
    By \cref{ax:zero}, we have $F^0_{\{1\}}(\delta_0) = \delta_1$, so $\gamma(0) = 0$.
    Note that $\Delta\{0,1\}$ is isomorphic to $[0,1]$, with $\delta_0$ and $\delta_1$ as the endpoints.
    From continuity of $F$ in $\beta$ (\cref{ax:diffble}) and the intermediate value theorem, for all distributions $p \in \mathrm{Int}\, \Delta\{0,1\}$, there is some $\hat\beta_p$ such that
    $F^{\hat\beta_p}_{\{1\}}(\delta_0) = p$, so for all $x \in [0,1)$, there is some $\beta_x$ such that $\gamma(\beta_x) = x$.
    % $\gamma(\beta_p) = p$
    Furthermore, $\gamma$ must be strictly increasing.
    If
    % $\beta > \beta'$
    $0 < x < y$ then again by the intermediate value theorem, there is some $\beta \in (0, \beta_y)$
    % such that $F^{\beta_q}_{\{1\}}(\delta_0)(X=1) $
    such that $\gamma(\beta) = x = \gamma(\beta_x)$.

    \TODO{}

    Thus $\gamma(t) $ is a continuous monotone function from $[0,\infty)$ to $[0,1]$ with $\gamma(0) = 0$,
    % (as $F^0_{\{1\}}(\delta_0)(X=1) = \delta_0(X=1) = 0$)
    and $\lim_{t\to\infty} \gamma(t) = \delta_1$.


    Conversely, given such a function, we can define an update rule for the singleton $A = \{1\}$, as
    \[
        F^{\beta}_{\{1\}}(p) :=
        \begin{bmatrix}
            1 \mapsto& \gamma(\beta - \beta_p) \\
            0 \mapsto& 1-\gamma(\beta - \beta_p)
        \end{bmatrix}.
    \]
    % \[
    %     F^{\beta}_{\{1\}}
    %
\end{lproof}

The selectors are of the form
\begin{align*}
    F_0^\beta(p) &:= \sigma( \sigma^{-1}(p) - \beta a)
\end{align*}
where $\sigma(x) = \frac{1}{1 + e^{-x}}$ is the logistic function,

\subsection{Finite Spaces}
Now, for some $n \in \mathbb N$, suppose $X := [n] = \{1, \ldots, n\}$, and $\mathcal A$ be all subsets of $X$.

\subsection{Gaussians}
\subsubsection{}
Consider the case where $\X = (\mathbb R \mathcal A)$ is the set of real numbers with the Borell $\sigma$-algebra, and $d(x,y) := (x-y)^2$.
Then the cannoncial update rule corresponds to multiplying by a Gaussian density, whose variance is $\nicefrac1\beta$.

Furthermore, applying it to a distribution which is already Gaussian with mean $\mu$ and variance $\sigma^2$, we find that
\begin{align*}
    F^{\beta}_{X=b}(\mathcal N(\mu, \sigma^2))(x) &\propto
        \exp\left\{ - \frac12 \frac{(x-\mu)^2}{ \sigma^2 } - \beta(x-b)^2\right\}
    \\&\propto \mathcal N(x| \mu', \sigma'^2)
\end{align*}
which is itself a Gaussian with mean equal to the weighted average of $\mu$ and $\beta$.

The upshot is that

\subsection{Population Dynamics}
Imagine that $\X$ is a space of possible phenotypes, and that a probability $\Pr \in \Delta\!\X$ represents a population distribution.

\subsubsection{Decay}
Let $\mathcal X = (X, \mathcal A)$ be a a measurable space.

% \[
%     F'(\mu)(x) :=   \ell(x) - \Ex\nolimits_\mu [ \ell ]
% \]


Now, consider a measurable lost function
$\ell: \X\times\mathcal A \to \mathbb R^+$.
We can now define the \emph{$\ell$-selector}, which is the
update rule governed by exponential decay, by:
% \def\cost#1#2{c(#1,#2)}
% \def\cost#1#2{d(#1,#2)}
\def\cost#1#2{\ell_{#2}(#1)}
\begin{align*}
    F^\beta_A(\Pr)(x) &\propto \Pr(x) \cdot \exp(-\beta \cost xA)
        \qquad\text{---or more formally,}\\
    F^\beta_A(\Pr)(B) &:=  \frac{1}{\Ex_{\Pr}[\exp(-\beta \cost xA )]}
        \int 1_B(x) \; \exp(-\beta \cost xA ) \,\mathrm d\Pr(x)
    %
\end{align*}

The name comes from the fact that, in our analogy to a population, an update of this kind corresponds to \emph{selection}.
% More concretely, applying $F^\beta_\ell$ to a population distributed according to $P$, is equivalent to the following algorithm:


% \begin{enumerate}[nosep]
%     \item Input: population distribution, a multiset $P$ of values of $x$, distributed according to $p(X)$.
%     \item Select a member $x$ of the population $P$.
%     \item
%         % Replace $x$ with $e^{-\ell(x)}$ copies of $x$.
%         Allow the population of $x$'s to grow/decay with rate $\ell(x)$, over a period $\beta$ of time.
%     \item Repeat.
% \end{enumerate}


% \begin{prop}
%     The cannonical update rule satisfies UR1-6.
% \end{prop}
%
%
% \begin{conj}
%     %[Characterization Theorem]
%     % Any update rule $F$ satisfying UR1-6 is a decay update rule for some cost function $c : \mathcal A \times X \to \mathbb R$.
%     Any update rule satisfying UR1-6 (++more) is a decay update rule for some cost function.
% \end{conj}


As a particular example, consider the loss $\ell_A(x) = \mathbbm1[x \in A]$.

It satisfies



\subsubsection{Flow}
\def\vgrad{\boldsymbol\nabla}
% \def\vgrad{\vec\nabla}
The update rule above only ``moves'' distributions by decay. If we imagine a distribution representing a population of individuals, the distribution has shifted by selection: an application of the update rule $F_A^\beta$ corresponds to culling the population, in proportion to their distance from $A$, over a time period of length $\beta$.
It is worth noting that such a population is then smaller, and must be re-normalized later.

A distribution could also shift by genuine motion of the underlying individuals.
In this case, let's model this as a partial differential equation in terms of $\Pr(x,t)$.
Now, we have some kind of conservation of mass, so from the continuity equation, we get
$\frac{\partial }{\partial t}Pr(x,t) = - \vgrad \cdot \mat J$,
where $\mat J(x,t)$ is the flux of individuals at point $x$ and time $t$.
Further assuming that the flux is generated by a combination of diffusion and a potential proportional to distance to $A$, we obtain

\[
    \frac{\partial \Pr(x, t)}{\partial t} = \vgrad_{\!x} \cdot \Big( k \vgrad_{\!x} \Pr (x) - \beta \vgrad_{\!x} d(x, A) \Big)
\]
where $k$ is a diffusion coefficient, and $\beta$

% To draw a different analogy, we can think of


\subsubsection{Drift}
We have talked about selection, but what about the second major force that drives evolution: random mutations?
% The important feature here is that not all information is truly used to find
% The result is that the population tends to ``spread out'', all else being equal,
% and has the effect of mixing the
One effect is that if there's no selection pressure to favor one variant over another, over time both variants will be represented equally.
Suppose that the elements of $\X$ have a canonical coded representation (such as DNA), such that specifying different phenotypes $x \in X$ requires different code lengths.
Because there is a tangible cost to carrying a larger genome, this induces a base measure with an inductive bias towards shorter codes.
% So the effect of random mutations is to mix

% \subsubsection{Sex}
% Sex is the thing that separates ``genetic algorithms'' from vanilla gradient descent.
%
% \begin{align*}
%     F^\beta_A(\Pr)(x) &\propto \Pr(x) \cdot \exp(-\beta \cost xA) \\
% \end{align*}

\subsection{Higher Order Probability Measures}
Now, suppose $\X$ is itself a set or probability measures.

\[
\]



Consider the following ways of extracting an (ordinary) probability measure $\mu(\Omega)$
from a higher order probability $\Pr(\mu(\Omega))$.

\begin{enumerate}
    \item \textbf{Centroid (Mean).}
        % $\mu^* := \textit{centroid}(\Pr)$.\\
        On a measurable set $A \in \mathcal A$, the centroid of $\Pr$ is defined as
        \[
            \mu^{\text{avg}}(A) := \Ex_{\mu \sim \Pr}[ \mu(A)] =
                \iint \Pr(\mu) \mu(x) 1_{A}(x)\, \mathrm d x\,\mathrm d\mu,
        \]
        the centroid is the usual ``flattening'' of a higher-order probability distribution.
        This is the ``multplication'' operation of the Giry Monad, the standard way in which a distribution over distributions is converted to a regular distribution.

    \item \textbf{Maximum Likelihood (Mode).}
        % $\textit{MLE}(\Pr) := A \mapsto (\arg\max_\mu \Pr(\mu))(A)$
        $\mathit{MLE}(\Pr) = \argmax_\mu \Pr(\mu)$
        This is the way that people often do inference.

    \item

    % \item \textbf{MAP Estimate.}
    %     Suppose we have a prior $\Pr_0$ over joint probabilities $\mu$. Then,
    %     \[
    %         \argmax_{\mu \in \Delta X} \Pr_0(\mu) \Pr(\mu)
    %     \]

\end{enumerate}


\subsection{PDGs}


\subsubsection{Updating Second-Order Distributions with PDGs}
Let $\mat X$ be a set of variables, where each variable $X \in \mat X$ can take possible values in the set $\V(X)$, and let $\V(\mat X)$ be the set of possible joint settings of all variables in $\mat X$.

\begin{align*}
    \Theta &:=
        \Big\{
        \text{Second-order probabilities}~ \Pr \in \Delta^2 \V(X)
        \Big\} \\
    % \Phi &:= \Big\{ \text{PDGs}~ \dg M
    %     \text{ with variables } \N^{\dg M} = \mat X \Big\}
    \Phi &:= \Big\{ \text{cpds}~ p(Y|X)
        % \text{ where }  X, Y \in \mat X \Big\}
        \text{ where }  X, Y \subset \mat X \Big\}
        %%% TODO: What about hypergraph fs non-hypergraph distinction?
\end{align*}

Now, consider the Bolzmann update rule
% Given a PDG $\dg M$, let $\X := \Delta\V(\dg M) = \Delta(\prod_{X\in\N}\V(X))$. For each edge we have
\[
    F_L^{\beta}(\Pr) (\mu) \propto \Pr(\mu) \exp(-\beta \kldiv\mu \bp)
\]

% Note, that the free extension of $\Phi$ is
Note, that the free extension of $\Phi$ is isomorphic to the set of ``purely quantitative'' PDGs, i.e., those with $\balpha = \mat 0$, over the variables $\mat X$:
$$
    % \bar\Phi = \{ \text{PDGs}~\dg M \text{ with variables } \N^{\dg M} = \mat X \Big\}
    \bar\Phi = \Big\{ \text{PDGs}~\dg M = (\mat X, \Ed, \bP[], \mat0, \bbeta) \Big\}
$$
and it is easily verified that update rule from before extends to
\[
    F_{\dg M}^k(\Pr)(\mu) \propto \Pr(\mu) \exp(- k \bbr{\dg M}_0(\mu))
\]
Note that
\begin{align*}
    \lim_{k \to \infty} F_{\dg M}^k(\Pr)(\mu) &\propto
        % \mathbbm1[\mu \in {\displaystyle\SD{\dg M}}]
        \Pr(\mu) \mathbbm1[\mu \in {\bbr{\dg M}^*_0}] \\
        &= \Pr \,|\, \bbr{\dg M}^*_0 \\
        &= \Pr \,|\, \SD{\dg M} \text{~if $\dg M$ is consistent.}
        % &= \mathrm{Unif}_{\{\!\!\{{\dg M}\}\!\!\}} \text{~if $\dg M$ is consistent.}
\end{align*}

That is to say, applying the update with observation $\dg M$ and infinite certainty is like conditioning the higher-order distribution $\Pr$ on those distributions $\mu$ consistent with $\dg M$.
% On the other hand,
Now we turn to the opposite extreme, when confidence is low. When $k=0$, clearly $F^k_\dg M(\Pr) = \Pr$, so $F$ satisfies \cref{ax:zero}. And in the limit, as $\gamma \to 0$,
\begin{align*}
    \lim_{k \to 0} F_{\dg M}^k(\Pr)(\mu) &= \Pr \,|\, \{ \Inc_{\dg M}(\mu) < \infty \} \\
        &= \Pr \,|\, \{\mu : \forall L.~\mu(Y|X) \ll \bp \}.
        % \quad\Big(\text{i.e., if $\mu$ is absolutely continuous with respect to $\bp$}
        % \Big)
\end{align*}

So $F$ can only be continuous at $k=0$ for $\Pr$ (and hence can only satisfy \cref{ax:diffble}) if, for all edges $L \in \Ed$, we have that $\Pr(\mu(Y|X) \ll \bp) = 1$.
So one easy way to ensure differentiability for all update rules is to require strictly positive probabilities in theprobabilities in the cpds.



\subsubsection{Updating Distributions with PDGs and Incompatibility}
% \def\tauur{\mathtt T}
% \def\tauur{\mathtt{GD}}
% \def\tauur{\tau}
\def\tauur{\mathtt{CPD\_UR}}

\begin{align*}
    \Theta &:=
        \Big\{
        \text{Probabilities}~ \mu \in \Delta \V(\mat X)
        \Big\} \\
    \Phi &:= \Big\{ \text{cpds}~ p(Y|X)
        % \text{ where }  X, Y \in \mat X \Big\}
        \text{ where }  X, Y \subset \mat X \Big\}
\end{align*}

Given a cpd $p(Y|X)$ with $X,Y \subset \mat X$,
define an update rule on joint distributions $\mu$ by
\begin{align*}
    % (\tau_{p(Y|X)}^\beta \mu) (\mat X) \propto
    %     \mu(X) \left(\frac{p(Y|X)}{\mu(Y|X)}\right)^{1 - e^{-\beta}} \mu(\mat X\setminus\{X,Y\}\mid X,Y)
    \Big(\tauur_{p(Y|X)}^\beta \mu\Big) (\mat X) &\propto
        \mu(\mat X) \left(\frac{p(Y|X)}{\mu(Y|X)}\right)^{1 - e^{-\beta}}
\end{align*}
    %
    % \\
    % &= \mu(X) p(Y|X)^{e^{-\beta}} \mu(Y|X)^{1-e^{-\beta}}  \mu(\mat X\setminus\{X,Y\}\mid X,Y)
%     &= \mu(X)\; p(Y|X)^{\epsilon} \mu(Y|X)^{1-\epsilon}  \mu(Z\mid X,Y)
% \end{align*}
Writing this in a
where $\epsilon := 1-e^{-\beta}$ and $Z := \mat X \setminus \{X,Y\}$.

\begin{prop}
    $\tauur_{p(Y|X)}$
\end{prop}

\begin{prop}
    The vector field of $\ext\tauur$ is the natural gradient of $\Inc_{\dg M}$, i.e.,
    $
        % \tau'_{\dg M}(\mu) = - \hat \nabla_\mu \bbr{\dg M}_0(\mu),
        \ext\tauur'_{\dg M} = - \hat \nabla \Inc_{\dg M},
    $
    so $\tauur$ is the update rule on joint distributions which reduces incompatibility by locally moving in the direction of steepest descent.
\end{prop}

\begin{prop}
    % \begin{enumerate}
        % \item \relax \texttt{\color{red}[CONJECTURE]}~
        % $\displaystyle
        %     \tauur_{\dg M}^{\{\nf1\gamma\}}(\mathrm{Unif}) \in \bbr{\dg M}_{\gamma}^*
        % $
        %
        % \item
        Updating $\dg M$ with absolute certainty computes the information projection onto the convex set of distributions consistent $\dg M$ i.e.,
        $$\displaystyle
            \tauur_{\dg M}^\infty(\mu) = \argmin_{\mu' \in \SD*{\dg M}} \kldiv{\mu'}{\mu}
        $$

    % \end{enumerate}
\end{prop}


\subsubsection{Updating PDGs with PDGs, via inconsistency reduction}
\begin{align*}
    \Theta :=
        \Big\{
        \text{PDGs}
        \Big\}; \qquad
    \Phi := \Big\{ \text{PDGs} \Big\}
\end{align*}

% Of course, we often think
We now consider the more general setting, in which both the observations $\Phi$ and the belief states $\Theta$ are
Here's an update rule that simply adds the new data to a PDG.
\[
    F^{\beta}_{p(Y|X)}(\dg M) = \dg M ~+~
        \begin{tikzpicture}[center base]
            \node[dpad0](X) {X};
            \node[dpad0,right=0.8 of X](Y) {Y};
            \draw[arr1](X) to
                node[above,inner sep=2pt]{$p$}
                node[below,inner sep=2pt]{${\scriptstyle\color{gray}(\beta)}$} (Y);
        \end{tikzpicture}
\]

It's extension is the update rule that incorporates data of the new PDG to the old one, where all confidences in the new PDG are scaled by $k$.
\[
    % F^{\beta}_{\dg M'}(\dg M) = \dg M ~+~ \beta\,\dg M'
    \bar F^{k}_{\dg M'}(\dg M) = \dg M ~+~ k\,\dg M'
\]

% There are actually several natural
The colletion of all PDGs, even on a finite set of finite variables, does not naturally form a manifold of fixed finite dimension.
But if we restrict our attention to PDGs with a certain fixed structure, then it does.
But, per \parencite{richardson2022loss}, we already have a natural choice of a loss $\mathcal L = \aar{-} : \mathrm{PDG} \times \mathrm{PDG} \to \mathbb R$: the inconsistency of the joint PDG.

Still, there is more than one reasonable way to perform updates: we can adjust the internal confidences $\bbeta$ of our PDG, or its cpds $\mat p$.
% The first is closer to a mixture of experts approach, and the latter

\[
    \bar G'_{\dg M'}(\dg M) = - \hat\nabla_{\mat p} \aar{\dg M + \dg M'}
\]

\begin{openQ}
    How do $\bar F$ and $\bar G$ relate to one another?
\end{openQ}

\begin{conj}
    $\tauur$ should be a special case.
\end{conj}

\subsubsection{}
\begin{align*}
    \Theta :=
        \Big\{
        \text{PDGs over variables } \mat X
        \Big\}; \qquad
    \Phi := \Big\{
        \text{Probabilities}~ \mu \in \Delta \V(\mat X)
        \Big\}
\end{align*}


\subsubsection*{The Qualitative Half}


There are a number of different ways to think about alpha.
\begin{enumerate}[nosep]
    \item $\alpha_L$ is your confidence in the functional dependence of $Y$ on $X$, given
    \item $\alpha_L$ is the ``alignment'' of the edge $L$
\end{enumerate}

We need a different distance function for the other half. Now we use
$d(\mu, DET)$.

% \begin{align*}
%
% \end{align*}

We need a different distance function for the qualitative loss function of a PDG.
Why measure distance differently in this case?


\section{Relationships to Other Notions of Confidence}

\subsection{}

Define an update rule on $\Phi = \mathcal A$ via its vector field,
\[
    F'_A(\mu) := [ \mathbbm 1_{A} ]
     \qquad = x \mapsto \mathbbm1\{x \in A\} - \mu(A)
\]
which results in
\[
    F_A^\beta(\mu)(x) \propto
    \begin{cases}
        \mu(x) e^{-\beta} & x \in A \\
        \mu(x) & x \notin A
    \end{cases}
\]
\[\text{? or}\quad
    F_A^\beta(\mu) (B) =
    % \frac{1}{\Ex_\mu[ e^{} ]}
    \frac{1}{1 + (e^{-\beta}-1) \mu(A) } \mu(B) e^{-\beta \mu(A \cap B)}
\]

\subsection{Probability Measures}

% Fix a probability $\Pr(X)$, and
We began by remarking that often confidence is taken to mean probability.
With this framework in mind, here is one explanation why.

Suppose $F$ is an update rule for assertions $\Phi_F = \mathcal A$.
Now consider a modified update rule $\tilde F$ over assertions $\Phi_{\tilde F} = \Delta\X$, given in vector field representation as
% Fix a probability distribution $p(X)$.
% consider the update rule $G = \mathdcal F_p[F]$ whose vector field is given by
\[
    \tilde F'_p(\mu) := \sum_{x} p(x) F'_{\{x\}}(\mu)
    = \Ex_{x \sim p} [ F'_{\{x\}} (\mu) ].
\]

\subsection{Belief Functions}
More genrally, for the space of assertions $\Phi$ consisting of mass functions $m \in \Delta(2^X)$ that assign each subset of X, encoding a belief function. Then

\[
    \tilde F'_p(\mu) := \sum_{A \in \mathcal A} m(A) F'_{A}(\mu).
\]

% Intuitively, $\mathdcal F_p$ is the rule
% A fixed point of $\mathdcal F_p$ corresponds to an update rule such
% One natural question: what update rules



\subsection{Epistemic Entrenchment}
\subsection{Weighted Probability Measures}

\section{}
% Objective:
Consider a density $\rho : \X \to \mathbb R^+$, and consider the differential equations
\[
    % \frac{1}{u(x,t)}
    \frac{\mathrm d}{\mathrm d t} u(x,t)
    = - \Big\langle\mathrm d\beta(t), \ell(x) \Big\rangle
        \,
        {u(x,t)}
    \qquad\qquad%% OR %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \dot {\mat u} = - \beta \mathrm{Diag}(\ell) \mat u
        % = - \Big\langle\mathrm d\beta(t), \ell(x) \Big\rangle
\]
Which means
\[
    % \Big|
        \log u (x,t)
    % \Big|
    = \int_{0}^{t}  \ell(x) \mathrm d \beta \mathrm d \tau
\]
% And suppose $C(t)$ is such that $u(x,t)$ is conserved, i.e.,





% \end{minipage}\label{ax:linear}\label{ax:linear}\label{ax:linear}\label{ax:linear}\label{ax:linear}\label{ax:linear}\label{ax:linear}





% What matters in the end are probabilities, as far as probabilistic models are concerned.
% But that doesn't mean that




% Probabilistic models may , but what matters in the end are probabilities.
In the opacity analogy, an update rule corresponds to a ``blending strategy'' --- a way of overlaying an image over another one, with a specified translucency.


\newpage
\printbibliography


\appendix
\section{General Thoughts}

\begin{itemize}
\item
(Variable) confidence is an important aspect of knowledge representation, because we need be able to entertain possibilities and reason about models we aren't sure about.

\item
Confidence is often thought of as the opposite of uncertainty.

\begin{itemize}
    \item
    Confidence is intimately related to probability, and our development will largely be couched in probabilistic terms. But they are not the same concept.
    While having high confidence in $\varphi$ is not so different from thinking $\varphi$ likely, having \emph{low confidence} in $\varphi$ is not the same as having thinking $\varphi$ is unlikely.
    To illustrate, consider a statement made by friend (a high-confidence source) and an adversary (a low-confidence source).

    \item
\end{itemize}

Rather, for us, it is \emph{conjugate} to uncertainty, in the way that force and distance are conjugate, and in the way that entropy and temperature are themodynamically conjugate variables.

Operationally, this means that it energetically unstable to have both high confidence and
% One has confidence in information.

\begin{wip}
! Why is $\mathit{Entropy} \cdot \mathit{Temperuature} = \mathit{Energy}$ but
 Why is $\mathit{Information} \cdot \mathit{Confidence} = \mathit{Energy}$ when confidence is like inverse temperature?
\end{wip}

\item
Confidence is, by nature, impossible to get an absolute scale for. This shows up again and again.
\begin{itemize}[nosep]
    \item PDG $\beta$ (and $\alpha$, relative to $\alpha_0$).
\end{itemize}

\item Confidence as inverse temperature.

\begin{itemize}
    \item
    ``cool-headed'' means calculated (a vote of confidence) while ``hot-headed'' means rash (such an assessment indicates a lack of confidence). Confidence is like inverse temperature.

    \item Boltzman rationality.
\end{itemize}


\item Confidence units: energy per inconsistency (at least when expressed as a gradient).

\item Pseudoinverse for constrained optimization
    \url{https://arxiv.org/pdf/1107.1944.pdf}

\item Geometrically, $\beta$ is the total distance to travel,

\item No Canonical Choice of Scale

\begin{itemize}
    \item Can choose choice of units to measure other things also (e.g., length in meters or feet).
\end{itemize}

\item Question:


\end{itemize}

% The notion of confidence also has
% Confidence is the opposite of uncertainty: you cannot be uncertain about $X$ .


% Confidence is dual to information.



Can you can be confident that $X$ is uncertain?
There is a huge difference between being certain that a coin has is fair, and knowing nothing about it.


% Here we draw a distinction between confidence and certainty:
% while you are certain that [proposition],




% Nevertheless, we submit that in many cases---and especially in more subjective or computationally restricted settings---probability alone is not enough, and the higher-order probabilistic picture is far too extreme.

Confidence as learning rate. Bigger confidence ~= bigger step size.
Learning algorithms have learning rate schedules (rates decrease or cycle).

\subsection*{Our Approach}
% Here is the idea: if you are confident in $X$, then you are
The idea is to view confidence as a property of an update, not as a property of a point of view.
In this telling

This allows us to distinguish between ...
%

% \begin{example}
% Let $X$ be a binary random variable, and suppose you have a prior probability $p$ that $X=1$.
% \begin{enumerate}
%     \item Suppose $p=1$.
%     This is a poor choice of prior.
%     If you then recieve information that $X=0$, something has gone very wrong, and the Bayesian update is undefined.
%
%     If you rceive information that $X = 1$, your internal state should not change.
%     % If we characterize
%
%     \item One might argue that only positive probabilities are relevant.
%     Suppose $p = 1 - \epsilon$.
% \end{enumerate}
% \end{example}


\subsection*{Issues To Address}
\begin{enumerate}
    \item The difference between having confidence in \emph{a source} and confidence in a particular \emph{fact}.  Is one more general than the other?
\end{enumerate}

Some intuitive features we might want to capture:
\begin{enumerate}

    \item \textbf{Trust Dynamics.} If a trusted source tells you something you already believe, your confidence in them goes up (maybe). Certainly if a source tells you something you know to be wrong, your confidence in the source goes down.  Whether or not you adopt the

    Also, if you ultimately end up adopting the belief, and later end upwith a more coherent picture of the world, your trust in the source should go way up.
    This is because in the end we want to place the most trust in sources that tell you the truth, not what you expect (or want) to hear.

    A source that tells you only exactly what you already believe ``artificially'' increases your confidence but does not actually provide you any information, unless they came to hold the same views independently.

\end{enumerate}

\TODO
\clearpage

\section{Useful Facts}
\begin{enumerate}
    \item Every length-minimizing curve in a Riemannian Manifold is a geodesic, if given a constant-speed parameterization
\end{enumerate}


\end{document}
